### 2024-02-05 presentations/grp.qmd
remove freeze

---
title: Presentation
format:
  html: default
  revealjs:
    output-file: 'grp-revealjs.html'
---

![](../figures/evolution/plasma-adiabatic-evolution-previous.svg)

![](../figures/evolution/plasma-adiabatic-evolution.svg)

![](../figures/evolution/plasma-adiabatic-evolution-next.svg)