### 2024-02-05 presentations/grp.qmd
update websites

---
title: Presentation
format:
  html: default
  revealjs:
    output-file: 'grp-revealjs.html'
code-links: 
  - href: ./grp-revealjs.html
    text: Slides
format-links: [revealjs]
---

## Plasma Adiabatic Evolution

![](../figures/evolution/plasma-adiabatic-evolution-previous.svg)

![](../figures/evolution/plasma-adiabatic-evolution.svg)

![](../figures/evolution/plasma-adiabatic-evolution-next.svg)