### 2024-02-11 notebooks/events/01.qmd
update

# 2019-04-07

``` txt
file:///Users/zijin/projects/autoplot/psp_mag_sw.vap?timerange=2019-04-01+through+2019-04-15
file:///Users/zijin/projects/autoplot/psp_mag_sw.vap?timerange=2019-04-06+through+2019-04-08
file:///Users/zijin/projects/autoplot/ace.vap?timerange=2019-04-08+through+2019-04-10
file:///Users/zijin/projects/autoplot/ace.vap?timerange=2019-04-11+through+2019-04-14
file:///Users/zijin/projects/autoplot/wind_plasma.vap?timerange=2019-04-08+through+2019-04-10
file:///Users/zijin/projects/autoplot/dscovr.vap?timerange=2019-04-08+through+2019-04-10
```

[ENLIL Solar Wind Prediction - Parker Solar Probe - Perihelion 02](http://helioweather.net/missions/psp/per02/)

::: {.column-page layout-nrow="1"}
{{< video http://helioweather.net/missions/psp/per02/anim/tuz-a7b1-d4t05x1-d200t1-donki_den1-psp-x15wifs3z1fo_med_201903-201904.mp4 >}}

{{< video http://helioweather.net/missions/psp/per02/anim/tuz-a7b1-d4t05x1-d200t1-donki_vel2e1-psp-x15ifs3z1fr_med_201903-201904.mp4 >}}
:::

![](../../images/paste-15.png)

![PSP Observation Overvall](../../images/paste-19.png)

::: {.column-page layout-nrow="1"}
![PSP observation](../../images/paste-18.png)

![ACE observation](../../images/paste-16.png)

![ACE observation (continued)](../../images/paste-17.png)
:::

Summary plot

::: {.column-screen-inset layout-nrow="1"}
![](https://psp-gateway.jhuapl.edu/backend/data/summary_plots/Summary_Plot_20190405.png)

![](https://psp-gateway.jhuapl.edu/backend/data/summary_plots/Summary_Plot_20190406.png)

![](https://psp-gateway.jhuapl.edu/backend/data/summary_plots/Summary_Plot_20190407.png)

![](https://psp-gateway.jhuapl.edu/backend/data/summary_plots/Summary_Plot_20190408.png)
:::