### 2024-02-11 notebooks/events/index.qmd

---
title: Events
listing: default
---

Starting point for the PSP encounter analysis.

-   @telloniSpacecraftRadialAlignments2023
-   @velliUnderstandingOriginsHeliosphere2020

[Orbit Plotter - PSP Science Gateway](https://psp-gateway.jhuapl.edu/website/Tools/OrbitPlotter)

## 2020-01-28 PSP-Earth, E4

-   [WHPI - Campaign: 2226](https://whpi.hao.ucar.edu/whpi_campaign-cr2226.php)
    -   [WHPI - Parker Perihelion 4 Campaign Highlights](https://whpi.hao.ucar.edu/whpi_campaign-cr2226-highlights.php)
-   [ENLIL Solar Wind Prediction - Parker Solar Probe - Perihelion 04](http://helioweather.net/missions/psp/per04/)

> The orbit of Parker Solar Probe (Parker) during the 4th encounter with the Sun presented a unique opportunity for a multi-mission and multi-observatory collaboration as the 4th perihelion occurred nearly at the ***Sun-Earth*** line. This unprecedented configuration allowed ground-based solar observatories to measure the solar surface that Parker was magnetically connected to while Parker simultaneously measured the near-Sun environment at the closest distance to the Sun recorded to date (27.8 RS). Additionally, near-Earth missions (e.g., ACE, ARTEMIS, and MMS) could then measure this solar wind once it reached 1 au to study the evolution of solar wind structures as they propagate radially outward from the Sun. Probes within the magnetosphere of Earth, and observatories focused on the ionosphere, could then study how these structures affect the near-Earth environment. To take advantage of this great opportunity for large-scale coordinated science, the Parker Project Science team and the Whole Heliosphere and Planetary Interactions (WHPI) group organized a campaign that has brought together observatories and missions from across the world. This page is meant to contain information and links relevant to the Parker Solar Probe Encounter 4 campaign. For more information, please contact Robert C. Allen (Robert.Allen\@jhuapl.edu), the Parker Project Science team coordinator for this effort.

Conjunction with In Situ Observations

> In addition to magnetic footpoint mapping for solar observers, ballistic mappings were performed to estimate when solar wind observed at Parker may reach 1 au at either STEREO-A or near-Earth observers. This shows that the footpoints of STEREO-A and Parker may have been near each other from Jan 26th, 2020 to Jan 28th, 2020 (where the red lines cross the black horizontal line at 0), while the footpoints of Earth and Parker may have been close *from Jan 29th, 2020 until Jan 31st, 2020* (where the blue lines cross the zero-line).

::: {.column-page layout-nrow="1"}
![](images/paste-1.png)

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E4_orbit.png)
:::

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/Lag_from_PSP_to_STA_and_Earth.png)

Start Date: 20200125

End Date: 20200131

``` txt
file:///Users/zijin/projects/autoplot/psp_mag_sw.vap?timerange=2020-01-20+through+2020-01-31
file:///Users/zijin/projects/autoplot/psp_mag_sw.vap?timerange=2020-01-27+through+2020-01-31
file:///Users/zijin/projects/autoplot/ace.vap?timerange=2019-04-08+through+2019-04-10
file:///Users/zijin/projects/autoplot/ace.vap?timerange=2019-04-11+through+2019-04-14
file:///Users/zijin/projects/autoplot/wind_plasma.vap?timerange=2019-04-08+through+2019-04-10
file:///Users/zijin/projects/autoplot/dscovr.vap?timerange=2019-04-08+through+2019-04-10
```

![](images/paste-2.png)

## ![](images/paste-5.png)![](images/paste-4.png)PSP Encounter Infos

[PSP Science Planning](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters)

{{< embed ../00_notes.ipynb#tbl-events >}}

## 2018-10-24 PSP Earth

```         
file:/Users/zijin/projects/psp_conjunction/files/psp_v.vap?timerange=2018-10-23+through+2018-10-25
```

PSP plasma data is very sparse (6 data point/day)

![](images/paste-13.png)

# Trange: Encounter 8

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_orbit.png) ![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_HEE.png)

> Using a nominal 400 km/s solar wind velocity, the following figures show the ballistic mapping in a helioprojective and heliographic frame, followed by an ecliptic view of the trajectory in the frame rotating with the Earth. Several Parker spirals are drawn, showing the possible magnetic connection to the Earth near May 9th.

> E8, C2, Q4, Q5, 24 April–1 May, 2021. Throughout this period SO will be in quadrature with Earth, appearing from Earth on the solar east limb. In rapid succession over E8, PSP will appear in quadrature from SO on the east limb (trailing), then in conjunction with SO at perihelion on 28 April, with PSP at a distance of 15.5RS, then in conjunction with STEREO-A on 29 April and with Earth on 30 April. STEREO-A will be positioned trailing Earth by 45◦ midway to SO in longitude. This PSP encounter is extremely favorable to a potential joint in situ–remote-sensing campaign comprising PSP, SO, STEREOA, SDO, Hinode, and ground-based instrumentation.

-   [WHPI - Campaign: PSP 8th Perhihelion](https://whpi.hao.ucar.edu/whpi_campaign-psp8p.php)
-   [WHPI - Campaign: PSP 15th Perhihelion](https://whpi.hao.ucar.edu/whpi_campaign-psp15p.php)

2021-05-09

05/09/2021

![](images/wind.png)

![](images/psp.png)

vap+cdaweb:ds=PSP_SWP_SPC_L3I&filter=psp&id=vp_moment_RTN\[:,0\];vp_moment_RTN\[:,1\];vp_moment_RTN\[:,2\]&timerange=2021-05-07+through+2021-05-08

\|rebundle(vp_moment_R,vp_moment_T,vp_moment_N)\|magnitude()

\|rebundle(psp_fld_l2_mag_RTN_1min\_\_0,psp_fld_l2_mag_RTN_1min\_\_0,psp_fld_l2_mag_RTN_1min\_\_0)\|magnitude()

-   ops: "\|rebundle(vp_moment_R,vp_moment_T,vp_moment_N)\|magnitude()" title: "PSP/SWP_SPC Proton bulk velocity from the 1st moment of the reduced distribution function in \[inertial\] RTN coordinate system (All Qualities)" y: label: V (km/s)
-   ops: "\|rebundle(B_R,B_T,B_N)\|magnitude()" y: label: B (nT) title: PSP_FLD/MAG_RTN_1min Magnetic field in RTN coordinates (1 minute cadence)

PSP/PSP/SWEAP/SPAN Partial Moment Velocity in RTN Coordinates and Sun reference frame \|rebundle(Vx RTN,Vy RTN,Vz RTN)\|magnitude()

Wind or ACE

![](images/paste-2.png)

![](images/paste-1.png)

<!-- ![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_1day_plots/2021/05/20210509-day.png) -->

::: {.column-page layout-nrow="1"}
![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2021/20210510-3day.png)

![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2021/20210513-3day.png)
:::

tr = '2021-05-06 22:46 to 2021-05-15 22:46' vp = getDataSet( 'vap+cdaweb:ds=PSP_SWP_SPC_L3I&filter=psp_SW&id=vp_moment_RTN', tr ) plot( 0, vp )

density5min= getDataSet( 'http://cdaweb.gsfc.nasa.gov/pub/data/ace/mag/level_2_cdaweb/mfi_k0/2017/ac_k0_mfi_20170117_v01.cdf?Magnitude') dst60min= getDataSet( 'http://cdaweb.gsfc.nasa.gov/pub/data/omni/omni_cdaweb/hourly/2017/omni2_h0_mrg1hr_20170101_v01.cdf?DST')

t5min= density5min.property(QDataSet.DEPEND_0)

( density, flux, dst ) = synchronize( t5min, density5min, flux4min, dst60min, nn=1 ) \# experimental function (note there is a built-in function in Autoplot v2017a. plot( 1, flux ) plot( 2, dst )

## Trange

> C4, 18 September, b 2021. This will be an inferior triple SO–PSP–STEREO-A conjunction, with SO inside the orbit of Venus around 0.6 AU, PSP at about 0.7 AU, and STEREO-A at 0.9 AU. This will be an occasion for joint in situ studies of the properties of turbulence and dynamics of a solar wind stream between 0.5 and 1 AU.

https://solar-mach.streamlit.app/?embedded=true&date=20210925&time=0000&coord_sys=0&plot_spirals=1&plot_sun_body_line=1&plot_trans=0&plot_markers=Numbers&long_offset=270&bodies=STEREO+A&bodies=Earth&bodies=BepiColombo&bodies=Parker+Solar+Probe&bodies=Solar+Orbiter&speeds=346&speeds=485&speeds=400&speeds=300&speeds=287&

![](images/paste-8.png)

![](images/paste-9.png)

![](images/paste-7.png)

![](images/paste-6.png)

![](images/paste-10.png)

::: {.column-page layout-nrow="1"}
![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2021/20210802-3day.png)

![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2021/20210805-3day.png)
:::

## Trange: Encounter 11, 2022-02

> E11, C6, Q8, Q9 23–27 February, 2022. PSP approaches encounter E11 beginning with a trailing quadrature Q8 on 23 February, 2022, appearing on the east limb of the Sun from SO, before rapidly going through conjunction with STEREO-A and then SO and Earth (24 and 25 February, respectively) at perihe- lion with the Earth–Sun–STEREO-A angle '40◦ and the Earth–

![](images/paste-3.png)

[![](images/paste-4.png)](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E11_HEE.png)

![](images/paste-5.png)

::: {.column-page layout-nrow="1"}
![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2022/20220225-3day.png)

![](https://www.ngdc.noaa.gov/dscovr/plots/dscovr_3day_plots/2022/20220228-3day.png)
:::