### 2024-02-29 article.qmd
update

---
title: Evolution of solar wind discontinuities in the inner heliosphere
subtitle: PSP and Earth conjunctions and alignments
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
affiliations:
  - id: ucla
    name: University of California, Los Angeles
---

![](figures/density_distribution){#fig-distribution}

![](figures/thickness_current){#fig-thickness-current}

::: {#fig-plasma-adiabatic-evolution .column-page}
![](figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR-next.svg)

![](figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR.svg)

<!-- Comparison of plasma dimensionless quantities (plasma beta and normalized plasma speed by Alfven speed) between PSP (red) and DSCOVR (blue). Yellow dots are predicted plasma beta and normalized speed using formulary from @perroneRadialEvolutionSolar2019. -->
:::

::: {#fig-examples .column-page}

![](figures/examples/psp_e2_example.png)

Solar wind discontinuities examples observed by PSP and ARTEMIS fitted by a logistic function in LMN coordinates.

In general discontinuities observed by PSP is much shorter (1~5 secs) with larger current density.
:::

![Comparision map of (normalized) thickness and (normalized) current density](figures/enc2/thickness_current_spc_n_spc.png){#fig-comp-map}

Solar wind properties

L:
j:
$\Delta v_l$