### 2024-05-08 notebooks/events/e11.qmd
update

# Encounter 11

-   Perihelion: 2022-02-25T15:38

> E11, C6, Q8, Q9 23–27 February, 2022. PSP approaches encounter E11 beginning with a trailing quadrature Q8 on 23 February, 2022, appearing on the east limb of the Sun from SO, before rapidly going through conjunction with STEREO-A and then SO and Earth (24 and 25 February, respectively) at perihe- lion with the Earth–Sun–STEREO-A angle '40◦ and the Earth– Sun–SO angle decreasing from '15◦ to 0◦ on 7 March, 2022, with SO at 0.5 AU. There will be clear opportunities here for cross calibration of magnetic ﬁeld measurements, as well as a true stereoscopic reconstruction of the solar corona crossed by PSP, with opportunities for all the PSP–SO objectives.

> The eleventh encounter of Parker Solar Probe provides observations down to 0.062 au (13.3 R~S~). Since the footpoints of Parker will be visible from Earth when it is near perihelion, the Parker team has again partnered with the Whole Heliosphere and Planetary Interactions (WHPI) group to coordinate remote observing during this encounter. The first footpoint prediction will be issued on February 18th followed by daily updates form February 21st through 28th, 2022. For more information, please contact Robert C. Allen ([Robert.Allen\@jhuapl.edu](mailto:Robert.Allen@jhuapl.edu)), the Parker Project Science team coordinator for this effort and visit the[WHPI page for this campaign](https://whpi.hao.ucar.edu/whpi_campaign-psp11p.php).

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E11_HEE.png)

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E11_orbit.png)

![](images/paste-14.png)

``` txt
file:///Users/zijin/projects/autoplot/scripts/psp_mag_sw.vap?timerange=2022-02-22+through+2022-03-01
file:///Users/zijin/projects/autoplot/scripts/ace.vap?timerange=2022-02-23+through+2022-03-02
```

------------------------------------------------------------------------

-   Perihelion: 2021-04-29
-   Start Date: 20210426 End Date: 20210503

-   [WHPI - Campaign: PSP 8th Perhihelion](https://whpi.hao.ucar.edu/whpi_campaign-psp8p.php)

> Using a nominal 400 km/s solar wind velocity, the following figures show the ballistic mapping in a helioprojective and heliographic frame, followed by an ecliptic view of the trajectory in the frame rotating with the Earth. Several Parker spirals are drawn, showing the possible magnetic connection to the Earth near May 9th.

> E8, C2, Q4, Q5, 24 April–1 May, 2021. Throughout this period SO will be in quadrature with Earth, appearing from Earth on the solar east limb. In rapid succession over E8, PSP will appear in quadrature from SO on the east limb (trailing), then in conjunction with SO at perihelion on 28 April, with PSP at a distance of 15.5RS, then in conjunction with STEREO-A on 29 April and with Earth on 30 April. STEREO-A will be positioned trailing Earth by 45◦ midway to SO in longitude. This PSP encounter is extremely favorable to a potential joint in situ–remote-sensing campaign comprising PSP, SO, STEREOA, SDO, Hinode, and ground-based instrumentation.

![](images/paste-10.png)

Start Date: 20210426 End Date: 20210503

``` txt
file:///Users/zijin/projects/autoplot/scripts/psp_mag_sw.vap?timerange=2021-04-26+through+2021-05-03
file:///Users/zijin/projects/autoplot/scripts/ace.vap?timerange=2021-04-27+through+2021-05-04
```

![](images/paste-11.png)

![](images/paste-12.png)

------------------------------------------------------------------------

```         
file:///Users/zijin/projects/autoplot/scripts/ace.vap?timerange=2019-04-11+through+2019-04-14
file:///Users/zijin/projects/autoplot/scripts/wind_plasma.vap?timerange=2019-04-08+through+2019-04-10
file:///Users/zijin/projects/autoplot/scripts/dscovr.vap?timerange=2019-04-08+through+2019-04-10
```

![](images/paste-7.png)

![](images/paste-9.png)