### 2024-05-12 notes.qmd
chore: update dates and move updates to a different file

# Updates

- [x] Why ARTEMIS electron data and Wind ion data?
- [ ] Disagreement between anisotropy observed and predicted by the theory
  - maybe just too localized anisotropy that is not well measured with available time resolution... 
  - or maybe more complex ion distribution, where nongyrotropy is more important than anisotropy
# TODO

- [ ] Plot Bl(t), jm(l) and ion anisotropy for these cases?

## 02/11/2024 (Encounter 4)

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc4/thickness_current)

![](figures/enc4/density_distribution)

![](figures/enc4/dvl)
:::

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc4/dn-dT)

![](figures/enc4/dT-dB)

![](figures/enc4/dn-dB)
:::

## 02/09/2024 (Encounter 2)

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc2/thickness_current)

![](figures/enc2/density_distribution)

![](figures/enc2/dvl)
:::

Pair plot

::: {.column-screen-inset layout-nrow="1"}

![](figures/enc2/dn-dT)

![](figures/enc2/dT-dB)

![](figures/enc2/dn-dB)
:::

##

![](images/paste-20.png)

Given timeseries $(R_a, \phi_a$, $R_b(t)$,

![](images/paste-11.png)

![](images/paste-12.png)