### 2024-05-12 presentations/rMeeting.qmd

---
title: "Current Sheets in the Solar Wind: Parker Solar Probe Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
format:
  html: default
  pptx:
    slide-level: 2
slide-number: true
date: "2024-05"
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms

See [sparthrb](./sparthb.qmd) for alignment and observation details
-->

## Examples of discontinuities

![](../figures/examples/psp_e2_example.png)

In general discontinuities observed by PSP are with

- shorter duration (1~5 secs)

  compared with 5-10 secs

- large current density (100-1000 nA/m^2)

  compared with 1-10 nA/m^2

## Identification/Detection method

![](../figures/examples/detection_liu2022.png)

@liuMagneticDiscontinuitiesSolar2022

Compared with adjacent intervals without sampling

## Properties of discontinuities

![](../figures/examples/psp_e2_example.png)

Every discontinuity's $l$ component is fitted by a $\tanh$ (logistic) function.

Combined with plasma data ($n$, $v$, $T$)

$L$: thickness of the discontinuity

$J_0$: maxium current density

---

If we expect the discontinuities to be generated locally, then the thickness of the discontinuities should be related to local plasma parameters:

- Normalize thickness by $d_i$ (ion inertial length) or $r_i$ (ion gyroradius)
- Normalize current density by $J_A$ (Alfven current density)

## Plasma adiabatic evolution

:::: {.columns}
::: {.column}
![](../figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR.svg)

![](../figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR-next.svg)

![](../figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR-previous.svg)
:::

::: {.column}

![](../images/sty3348fig5.jpeg)

@perroneRadialEvolutionSolar2019
:::
::::

<!-- 
The radial dependence of the proton number density, $n_p$ and magnetic ﬁeld magnitude, $B$, is given by

$$
n_p = (2.4 ± 0.1)(R/R_0)^{−(1.96±0.07)} cm^{−3}
$$

$$
B = (5.7 ± 0.2)(R/R_0)^{−(1.59±0.06)} nT
$$

The faster decrease of the magnetic than kinetic pressure is reﬂected in the radial proton plasma beta variation

$$
β_p = P_k/P_B = (0.55 ± 0.04)(R/R_0)^{(0.4±0.1)}
$$ -->

## Comparison with discontinuities properties

![](../figures/enc2/thickness_current.spi_n_spi.png)

![](../figures/enc2/density_distribution.spi_n_spi.png)

K1: Normalized discontinuities properties does not vary much with radial distance

---

![](../figures/enc2/thickness_current.spc_n_spc.png)

![](../figures/enc2/density_distribution.spc_n_spc.png)

K2: $Δv_l$,l demonstrates a strong correlation with $Δv_{A}$, albeit being consistently smaller.

## Comparison with Alfven speed and plasma speed change

For rotational discontinuities, the plasma velocity jump across them: the plasma ﬂow velocity $v_l$ (the solar wind velocity projected onto $l$) changes due to changes in $B_l$, $Δv_l=±Δv_A$ with $v_A= B_l / \sqrt{4 n m}$ [@hudsonDiscontinuitiesAnisotropicPlasma1970].

![](../figures/enc2/dvl.spi_n_spi.png)

## Comparison with anisotropy

The equation for $Δv_A$ includes a factor depending on the anisotropy of the plasma.

![](../figures/enc2/anisotropy.spi_n_spi.png)

Pressure anisotropy $Λ = \mu (P_∥ - P_⊥) / B^2$.

## Conclusion

- The properties of the discontinuities is related to the local plasma parameters
  - Thickness => ion inertial length
  - Current density => Alfven velocity (current density)
- Normalized thickness and current density of discontinuities remain constant with radial distance
- Anisotropy of the plasma is expected to be larger near the sun to explain the observed speed change ratio $Δv_i/Δv_A$.

---

## References