### 2024-05-13 presentations/rMeeting.qmd
chore: move content locations

---
title: "Current Sheets in the Solar Wind: Parker Solar Probe Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
format:
  html: default
  pptx:
    slide-level: 2
slide-number: true
date: "2024-05"
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms
See [sparthrb](./sparthb.qmd) for alignment and observation details
See [Main Website](../index.qmd) for the results
-->