### 2024-08-28 presentations/rMeeting.qmd
chore: update website

---
title: "Current Sheets in the Solar Wind: Parker Solar Probe Observations"
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
date: "2024-05"
---

<!-- 
See [share notes](../../beforerr/docs/research/topics/swd/index.qmd) for
- introduction
- formation mechanisms
See [sparthrb](./sparthb.qmd) for alignment and observation details
See [Main Website](../index.qmd) for the results
-->