### 2024-11-04 notebooks/11_alignment.qmd
update: new section on alignment effects and discontinuity properties

## Effect of alignment on discontinuity properties

```{julia}
include("../src/main.jl")
using CairoMakie
using Beforerr

df_enc_7 = load_all(7);
df_enc_8 = load_all(8);
df_enc_9 = load_all(9);
df = reduce(vcat, [df_enc_7, df_enc_8, df_enc_9], cols=:intersect);
```

```{julia}
layer = data(df) * mapping(color=:dataset, col=enc_m)

function plot_p1()
    maps = [l_log_map, l_norm_log_map, j_log_map, j_norm_log_map]
    figure = (size=(450, 800),)
    f = plot_dist(layer, maps; figure, axis)

    # fine tune the plot
    for i in 1:4
        axs = contents(f[i, 1][:, :])
        ylims!.(axs; low=1e-3)
    end
    axs = contents(f[1, 1][:, :])
    xlims!.(axs; low=-1)
    axs = contents(f[2, 1][:, :])
    xlims!.(axs; low=-1.5)
    axs = contents(f[4, 1][:, :])
    xlims!.(axs; high=1)

    easy_save("thickness_current")
end
```

![](../figures/thickness_current.png)

```{julia}
function plot_B_angle(; figure=(size=(450, 800),), name = "B_angle")
    maps = [M.dB_norm_over_B, M.bn, ω_map, M.ω_in]
    f = plot_dist(layer, maps; figure)

    # fine tune the plot
    gl3_axs = contents(f[3, 1][:, :])
    gl4_axs = contents(f[4, 1][:, :])
    foreach(vcat(gl3_axs, gl4_axs)) do ax
        vlines!(ax, [30]; color=:gray, linestyle=:dash)
    end

    easy_save(name)
end

f = plot_B_angle()
```

![](../figures/B_angle.png)