### 2025-02-18 notebooks/10_results.qmd
docs: update

---
title: Results
engine: julia
---

```{julia}
using Revise
using DrWatson
@quickactivate
include(srcdir("main.jl"))
import Beforerr: easy_save
```

```{julia}
enc = 8 # event id (encounter id)    
easy_save(fname; enc=enc) = Beforerr.easy_save(fname; dir="$fig_dir/enc$enc", formats=[:svg, :pdf])
```

```{julia}
df_enc_7 = load_all(7);
df_enc_8 = load_all(8);
df_enc_9 = load_all(9);
df = reduce(vcat, [df_enc_7, df_enc_8, df_enc_9], cols=:intersect);
```

```{julia}
function plot_dist_comp(
    df1, df2;
    figure_kwargs=(size=(800, 600),),
    kwargs...
)

    df = reduce(vcat, [df1, df2], cols=:intersect)
    data_layer = data(df) * mapping(color=ds_mapping, marker=ds_mapping)

    maps = [l_log_map j_log_map
        l_norm_log_map j_norm_log_map]

    plot_dist(data_layer; maps, axis=(), visual=mapping(), figure_kwargs=figure_kwargs, kwargs...)
end

function plot_dist_comp(df1, df2, tr1, tr2)
    df1 = subset_timerange(df1, tr1...)
    df2 = subset_timerange(df2, tr2...)
    @info "Number of data points in df1: $(nrow(df1))"
    @info "Number of data points in df2: $(nrow(df2))"
    plot_dist_comp(df1, df2)
end
```

## Distribution comparison

### Enc 7

```{julia}
enc = 7 # event id (encounter id)    
easy_save(fname) = Beforerr.easy_save(fname; dir="$fig_dir/enc$enc", formats=[:svg])
mkpath("$fig_dir/$enc");
```

```{julia}
plot_dist_comp(psp_df, wind_df,
    ["2021-01-15", "2021-01-17"],
    ["2021-01-17", "2021-01-19"]
)
```

```{julia}
plot_dist_comp(psp_df, wind_df,
    ["2021-01-15", "2021-01-17"],
    ["2021-01-21", "2021-01-23"]
)
```

### Enc 8

```{julia}
plot_dist_comp(psp_df, solo_df)
```

```{julia}
DateTime("2021-04-29T00:45") + Hour(145)
DateTime("2021-04-29T00:45") + Hour(-1)
```

```{julia}
#| layout-ncol: 3
#| column: screen
f1 = plot_dist_comp(psp_df, solo_df,
    ["2021-04-29T00:15", "2021-04-29T02:15"],
    ["2021-05-04T17:45:00", "2021-05-04T19:45:00"]
)
f2 = plot_dist_comp(psp_df, solo_df,
    ["2021-04-28T00:15", "2021-04-29T02:15"],
    ["2021-05-03T17:45:00", "2021-05-04T19:45:00"]
)

f3 = plot_dist_comp(psp_df, solo_df,
    ["2021-04-28T00:15", "2021-04-29T02:15"],
    ["2021-05-04T17:45:00", "2021-05-05T19:45:00"]
)
display(f1)
display(f2)
display(f3);
```

```{julia}
begin
    df = reduce(
        vcat,
        [psp_df, thm_df, wind_df, solo_df],
        cols=:intersect
    )
    levels!(df.dataset, ds_order)

    earth_df = reduce(
        vcat,
        [thm_df, wind_df],
        cols=:intersect
    )
end;
```

```{julia}
begin
    # base data layer
    data_layer = data(df) * mapping(color=ds_mapping, marker=ds_mapping)

    # data layer with different alpha
    data_layer_a = (
        (data(thm_df) + data(wind_df)) * visual(alpha=0.3) + data(psp_df) * visual(alpha=1)
    ) * mapping(
        color=ds_mapping, marker=ds_mapping
    )

    # data layer with 1au missions
    data_layer_b = (
        (data(wind_df) + data(thm_df))
    ) * mapping(
        color=:dataset, marker=:dataset
    )
end;
```

## Discontinuities' properties

```{julia}
# velocity v is normalized to $l Ω$, where $l$ represents the width of the RD and $Ω = q B/(mpc)$ is the cyclotron frequency.
function v_norm!(df::AbstractDataFrame)
    df.v_to_norm = df.L_k .* df."B.mean"
end
v_norm!(df);
```

```{julia}
layer = data(df) * mapping(color=:dataset, col=:enc)
axis = (yscale=log10,)
figure = (size=(600, 800),);
```

```{julia}
Bn = :bn_over_b => abs => L"B_N/B";
v_norm = :v_to_norm => log10 => "Log v_gyro";
```

```{julia}
dfplot(layer, args...; axis=(;), kwargs...) = draw(layer * mapping(args...; kwargs...); axis)
dfplot(df::AbstractDataFrame, args...; axis=(;), kwargs...) = draw(data(df) * mapping(args...; kwargs...); axis)
```

```{julia}
using GLMakie
GLMakie.activate!()
axis = (type=Axis3, width=300, height=300)
dfplot(layer, v_norm, ω_in, Bn; axis)
```

```{julia}
maps = [B_log_map, v_norm, ω_in]
plot_dist(layer, maps; figure=figure)
```

```{julia}
function plot_dB_over_B(layer; axis=(yscale=identity,), datalimits=((0, 0.3),))
    plt = layer * mapping(dB_over_B_map) * density(datalimits=datalimits) * visual(Lines)
    draw(plt; axis=axis)
end

# log scale is not good for this parameter
# plot_dB_over_B(layer; axis = (yscale=log10,)) |> display

plot_dB_over_B(layer)
```

```{julia}
# plot the density distribution of the thickness and current density
begin
    fname = "properties_distribution"

    maps = [l_log_map j_log_map
        l_norm_log_map j_norm_log_map]
    figure_kwargs = (size=(800, 600),)

    plot_dist(data_layer; maps, axis=(), visual=mapping(), figure_kwargs=figure_kwargs)
    easy_save(fname)
end
```

```{julia}
function plot_l_j_local_comp(df)
    fig = Figure(size=(1000, 800))

    data_layer = data(df) * mapping(col=:dataset)

    facet = (; linkxaxes=:none, linkyaxes=:none)
    # Thickness Panels
    begin
        plt = data_layer * mapping(di_log_map, l_log_map)
        axis = (;)
        draw!(fig[1, 1:3], plt, axis=axis, facet=facet)
    end

    # Current Density Panels
    begin
        plt = data_layer * mapping(jA_log_map, j_log_map)
        axis = (;)
        draw!(fig[2, 1:3], plt, axis=axis, facet=facet)
    end

    fig
end

df |> plot_l_j_local_comp
```

```{julia}
begin
    fname = "thickness_current"

    plt = data_layer_a * visual(Scatter)

    fig = Figure(size=(1000, 500))
    axis = log_axis

    ga = fig[1, 1]
    gb = fig[1, 2]

    grid1 = plt * mapping(l_map, j_map) |> draw!(ga; axis=axis)
    grid2 = plt * mapping(l_norm_map, j_norm_map) |> draw!(gb; axis=axis)

    pretty_legend!(fig, grid1)
    add_labels!([ga, gb])
    easy_save(fname)
end
```

`ggscatterhist` can not handle log scale well

```{julia}
#| label: fig-dn-dB
#| column: screen
axis = (limits=((-0.5, 0.5), (-0.5, 0.5)),)

begin
    fname = "dn-dB"
    plt = data_layer_a * mapping(dn_over_n, dB_over_B) * (AlgebraOfGraphics.linear() + visual(Scatter))
    fg = draw(plt, axis=axis)
    easy_save(fname)
end
```

```{julia}
#| label: fig-dn-dT
#| column: screen
begin
    fname = "dn-dT"
    plt = data_layer_a * mapping(dT_over_T, dn_over_n) * (mapping())

    fig = Figure(size=(1000, 500))
    grid1 = plt |> draw!(fig[1, 1])
    grid2 = plt |> draw!(fig[1, 2]; axis=(; limits=((-0.5, 0.5), (-0.5, 0.5))))

    pretty_legend!(fig, grid1)
    easy_save(fname)

    fig
end
```

```{julia}
#| label: fig-dT-dB
#| column: screen
begin
    fname = "dT-dB"
    plt = data_layer_a * mapping(dT_over_T, dB_over_B) * (AlgebraOfGraphics.linear() + mapping())

    fig = Figure(size=(1000, 500))
    grid = draw!(fig[1, 1], plt)
    draw!(fig[1, 2], plt, axis=(; limits=((-0.5, 0.5), (-0.5, 0.5))))

    pretty_legend!(fig, grid1)
    easy_save(fname)

    fig
end
```

```{julia}
ω_map = :rotation_angle => "rotation angle (deg)";
v_l_ratio_map = :v_l_ratio => L"Δ V_{ion,l} / Δ V_{A,l}"
plt = data(psp_df) * mapping(ω_map, v_l_ratio_map) * (linear() + visual(Scatter))
draw(plt)
```

## Alfvenicity

Comparison of change in Alfven speed and plasma speed

The distinguishing feature of rotational discontinuities is the plasma velocity jump across them: the plasma ﬂow velocity $v_l$ (the solar wind velocity projected onto l) changes due to changes in $B_l$, $Δv_l=±Δv_A$ with $v_A= B_l / \sqrt{4 n m}$ (Landau & Lifshitz 1960; Hudson 1970).

```{julia}
#| label: fig-dvl
#| column: screen
fig = plot_dvl(psp_df)
```

```{julia}
plot_vl_ratio(df)
```

```{julia}
vl_ratio_label = L"\Delta V_{i,l} / \Delta V_{A,l}"
v_l_fit_ratio_map = "v_l_fit_ratio" => vl_ratio_label

plot_vl_ratio(psp_df; mapping_args=(v_l_fit_ratio_map,), fname=missing)
```

```{julia}
v_l_fit_ratio_map = "v_l_ratio.fit.Λ_ion" => vl_ratio_label

plot_vl_ratio(psp_df; mapping_args=(v_l_fit_ratio_map,), fname=missing)
```

## Anistropy

@shenComparingPlasmaAnisotropy2024

```{julia}
plot_anistropy_comparison_2d(psp_df)
```

```{julia}
psp_df.Λ_ion_r = abs.(psp_df.Λ_ion ./ psp_df.Λ_t .- 1)
first(sort(psp_df, :Λ_ion_r), 4)
```

```{julia}
plot_anistropy(df)
```

```{julia}
#| label: fig-anisotropy
#| column: screen
plot_anistropy_theory(df)
```

```{julia}
function plot_anistropy_comparison()
    fname = "anisotropy_comparison"

    fig = Figure(size=(1000, 800))

    plt_Λ = data(
        stack(earth_df, [:Λ, :Λ_t, :Λ_ion, :Λ_e], [:time, :dataset])
    ) * mapping(:value => L"\Lambda", color=:variable, marker=:variable, col=:dataset)

    grid1 = plt_Λ * density() |> draw!$fig[1, 1:2]
    grid2 = plt_Λ * histogram() |> draw!$fig[2, 1:2]
    # easy_save("$(fname)_$psp_p_instr")
    # legend!(fig[1, 1:2], grid, titleposition=:left, orientation=:horizontal)
    pretty_legend!(fig, grid1)

    fig
end

function plot_anistropy_comparison_i_e()
    fname = "anisotropy_comparison_i_e"

    mapping_layer = mapping(
        Λ_e_map, Λ_ion_map,
    )

    # draw a dashed line with slope 1
    df2 = (Λ_e=[0, 0.3], Λ_ion=[0, 0.3])

    plt = (data(earth_df) * mapping(row=:dataset) + data(df2) * visual(Lines)) * mapping_layer

    fig = draw(plt, facet=(; linkxaxes=:minimal, linkyaxes=:minimal))
    fig
end
```

```{julia}
#| label: fig-anisotropy-comp
#| column: screen
plot_anistropy_comparison()
```

```{julia}
#| label: fig-anisotropy-comp-2d
#| column: screen
plot_anistropy_comparison_2d()
```

```{julia}
#| label: fig-anisotropy-comp-i-e
#| column: screen
plot_anistropy_comparison_i_e()
```