### 2025-06-12 notebooks/000_data.qmd

## Data

- [PSP/FIELDS Level 2 and higher data](https://fields.ssl.berkeley.edu/data/), [SPDF - Coordinated Data Analysis Web (CDAWeb)](https://cdaweb.gsfc.nasa.gov/cgi-bin/eval2.cgi?dataset=PSP_FLD_L2_MAG_RTN&index=sp_phys)

Notes: `thb_peim_t3_magQ` is not accurate in solar wind mode.

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SPEDAS: tplot
using GLMakie
```

### THEMIS

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        THEMIS.B_GSE,
        THEMIS.B_FGL_GSE,
        THEMIS.n,
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

### Wind

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        [Wind.V_GSE_3DP, Wind.V_GSE_K0],
        [Wind.n_p_K0, Wind.n_p_3DP, Wind.n_p_nonlin],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

`Wind.V_GSE_3DP` ("WI_PM_3DP/P_VELS") has higer time resolution than `Wind.V_GSE_K0` ("WI_K0_SWE/V_GSE").

`Wind.n_p_nonlin` and `Wind.n_p_K0` seem to be the same.

![](../figures/others/wind.png)

WI_H1_SWE	Proton_W_moment	        Proton thermal speed W (km/s) from isotropic moment analysis
WI_H1_SWE	Proton_Wperp_moment 	Proton thermal speed Wperpendicular (km/s) from bimax moment analysis
WI_H1_SWE	Proton_Wpar_moment	    Proton thermal speed Wparallel (km/s) from bimax moment analysis

WI_H1_SWE	Proton_Wperp_nonlin	    Proton thermal speed perpendicular to the magnetic field direction [km/s].
WI_H1_SWE	Proton_Wpar_nonlin	    Proton thermal speed parallel to the magnetic field direction [km/s].
WI_PLSP_3DP	Ion omnidirectional fluxes 0.4-3 keV and moments, often at ~24 second resolution, PESA Low, Wind 3DP
WI_PM_3DP	Ion moments (computed on-board) @ 3 second (spin) resolution, PESA LOW, Wind 3DP