### 2025-06-29 notebooks/01_flux.qmd
feat: flux

```{julia}
using Unitful
using UnitfulAstro
using Unitful.DefaultSymbols
using Speasy

r = get_data("cda/PSP_COHO1HR_MERGED_MAG_PLASMA/radialDistance", "2022-12-11", "2022-12-12")
```

```{julia}
# Constants
using Unitful: k, G, μ0, mp
using UnitfulAstro: Msun
const mp = 1.6726e-27u"kg"

"""
Calculate fluxes given parameters at a spacecraft
"""
function compute_fluxes(n, V, T, B, r, Ω=1)
    Br = B[1]
    Vr = V[1]
    A = Ω * r^2 # flux tube cross-sectional area

    # Mass flux
    F_mass = n * mp * Vr * A

    # Magnetic flux (radial)
    F_mag = Br * A

    # Energy flux components
    E_kin = 0.5 * ρ * V^2 * Vr
    E_th = (5 / 2) * u * n * k * T
    E_grav = -ρ * u * (G * Msun / r)
    E_mag = (B^2 / (2μ₀)) * (u / ρ)

    # Total energy flux
    E_total = (E_kin + E_th + E_grav + E_mag) * A

    return (mass_flux=F_mass, mag_flux=F_mag,
        E_kin=E_kin * A, E_th=E_th * A, E_grav=E_grav * A, E_mag=E_mag * A,
        E_total=E_total)
end
```