### 2025-07-27 notebooks/000_data.qmd
update

## Data

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SPEDAS: tplot, set_meta
using GLMakie
```

### Parker Solar Probe

- [PSP/FIELDS Level 2 and higher data](https://fields.ssl.berkeley.edu/data/), [SPDF - Coordinated Data Analysis Web (CDAWeb)](https://cdaweb.gsfc.nasa.gov/cgi-bin/eval2.cgi?dataset=PSP_FLD_L2_MAG_RTN&index=sp_phys)

The parameters are:

- 'psp_fld_l2_mag_RTN_4_Sa_per_Cyc' (B in RTN, 0.25s resolution)
- 'psp_swp_spi_sf00_L3_DENS' (Ion density, regular resolution)
- 'psp_swp_spi_af00_L3_DENS' (Ion density, burst resolution)
- 'psp_swp_spi_sf00_L3_VEL_RTN_SUN' (velocity in RTN in Sun's frame, regular resolution)
- 'psp_swp_spi_af00_L3_VEL_RTN_SUN' (velocity in RTN in Sun's frame, burst resolution)
- 'psp_swp_spi_sf00_L3_VEL_RTN' (velocity in RTN in spacecraft frame, regular resolution, you need this to calculate current sheet thickness)
- 'psp_swp_spi_af00_L3_VEL_RTN' (velocity in RTN in spacecraft frame, burst resolution, you need this to calculate current sheet thickness)
- 'Tp_spani_b' (Ion parallel and perpendicular ion temperature, regular resolution. Blue (first element in array): parallel, Red: perp)
- 'Tp_spanib_b' (Ion parallel and perpendicular ion temperature, burst resolution. Blue (first element in array): parallel, Red: perp)

the helium-to-proton abundance ratio ($A_{He} = N_α/N_p × 100\%$, where $N_α$ and $N_p$ are the number densities of helium particles and protons, respectively)

```{julia}
tmin = DateTime("2021-01-15")
n_p = PSP.n_spi(tmin, tmin + Day(1))
n_α = PSP.n_alpha(tmin, tmin + Day(1))

A_He = set_meta(n_α ./ n_p .* 100; ylabel="A_He", units="%", scale=identity)

let tmin = DateTime("2021-01-15")
    tvars = (
        PSP.n_spi,
        PSP.n_alpha,
        A_He
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

### THEMIS

Notes: `thb_peim_t3_magQ` is not accurate in solar wind mode.

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        THEMIS.B_GSE,
        THEMIS.B_FGL_GSE,
        THEMIS.n,
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

### Wind

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
    # [Wind.V_GSE_3DP, Wind.V_GSE_K0],
        [
        unitify("1/cm^3") ∘ SpeasyProduct("WI_K0_SWE/Np"; labels=["Proton"]),
        SpeasyProduct("WI_PM_3DP/P_DENS"; labels=["Proton (3DP, PESA Low)"]),
        SpeasyProduct("WI_H1_SWE/Proton_Np_nonlin", labels=["Proton (non-linear fitting)"]),
        SpeasyProduct("WI_H1_SWE/Alpha_Na_nonlin", labels=["Alpha (non-linear fitting)"]),
        Wind.n_e
    ],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

`Wind.V_GSE_3DP` ("WI_PM_3DP/P_VELS") has higer time resolution than `Wind.V_GSE_K0` ("WI_K0_SWE/V_GSE").

`Wind.n_p_nonlin` and `Wind.n_p_K0` seem to be the same.

![](../figures/others/wind.png)

WI_H1_SWE	Proton_W_moment	        Proton thermal speed W (km/s) from isotropic moment analysis
WI_H1_SWE	Proton_Wperp_moment 	Proton thermal speed Wperpendicular (km/s) from bimax moment analysis
WI_H1_SWE	Proton_Wpar_moment	    Proton thermal speed Wparallel (km/s) from bimax moment analysis

WI_H1_SWE	Proton_Wperp_nonlin	    Proton thermal speed perpendicular to the magnetic field direction [km/s].
WI_H1_SWE	Proton_Wpar_nonlin	    Proton thermal speed parallel to the magnetic field direction [km/s].
WI_PLSP_3DP	Ion omnidirectional fluxes 0.4-3 keV and moments, often at ~24 second resolution, PESA Low, Wind 3DP
WI_PM_3DP	Ion moments (computed on-board) @ 3 second (spin) resolution, PESA LOW, Wind 3DP

### SOLO

"cda/SOLO_L2_MAG-RTN-NORMAL-1-MINUTE/B_RTN",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/N",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/V_RTN",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/T",

### STEREO

"cda/STA_L2_PLA_1DMAX_1MIN/proton_number_density",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_bulk_speed",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_n_s_flow_angle_inst",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_e_w_flow_angle_inst",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_Vr_RTN",