### 2025-07-27 notebooks/11_spectra.qmd

<a href="https://docs.juliadsp.org/stable/periodograms/">Periodograms - periodogram estimation · DSP.jl</a>

<a href="https://github.com/lootie/Multitaper.jl">lootie/Multitaper.jl: A Julia package for frequency domain analysis of time series</a>

[pmtm - Multitaper power spectral density estimate - MATLAB](https://www.mathworks.com/help/signal/ref/pmtm.html)

Example spectra of Ev, Eb, and Er, calculated over a 2 day interval (2007 July 6 00:00:00–2007 July 8 00:00:00) using the kinetic normalization, are shown in Figure 1.

```{julia}
using Dates
using PSPAnisotropy: Wind
tmin = DateTime("2007-07-06")
tmax = DateTime("2007-07-08")
B = Wind.B(tmin, tmax; add_unit=false)
```

```{julia}
using DSP

pxx = periodogram(B[:, 1])

```

```{julia}
# first, add the package if you haven't already:
# import Pkg; Pkg.add("Multitaper")

using Statistics
using LinearAlgebra
using Multitaper

const μ0 = 4π * 1e-7     # vacuum permeability [H/m]

"""
    compute_fluctuations(v, B, rho)

Given
  • v    : N×3 array of velocity (m/s)
  • B    : N×3 array of magnetic field (T)
  • rho  : length‐N vector of mass density (kg/m³)

Returns
  • δv   : N×3 array of velocity fluctuations
  • δb   : N×3 array of magnetic‐field fluctuations in Alfvén units (m/s)
"""
function compute_fluctuations(v::AbstractMatrix, B::AbstractMatrix, rho::AbstractVector)
    @assert size(v) == size(B) "v and B must be the same size"
    @assert size(v, 1) == length(rho) "rho length must match rows of v,B"

    # subtract mean
    v̄ = mean(v, dims=1)
    δv = v .- v̄

    # magnetic field in velocity units: b = B / sqrt(μ0 ρ)
    b = B ./ sqrt.(μ0 .* rho)[:, None]
    b̄ = mean(b, dims=1)
    δb = b .- b̄

    return δv, δb
end

"""
    trace_psd_mt(x, dt; NW=4)

Compute the trace (sum over 3 components) PSD of a 3‐component fluctuation array x via multitaper.

Arguments:
  • x    : N×3 array of fluctuation time series
  • dt   : sampling interval [s]
Keyword:
  • NW   : time–bandwidth product (default 4)

Returns
  • f    : vector of frequency bins [Hz]
  • P    : vector of trace power spectrum P(f) = ∑ᵢ S_ii(f)
"""
function trace_psd_mt(x::AbstractMatrix, dt::Real; NW::Real=4)
    N, nc = size(x)
    @assert nc == 3 "x must be N×3"

    fs = 1 / dt
    # how many tapers? standard choice K = 2NW − 1
    K = Int(2 * NW - 1)

    # accumulate over components
    f = nothing
    P_trace = nothing

    for comp in 1:3
        # multitaper spectrum for this component
        spec = mt_spectrum(x[:, comp];
            NW=NW,
            K=K,
            fs=fs,
            adaptive=false)  # or true for adaptive weighting

        f === nothing && (f = spec.f)
        P_trace === nothing && (P_trace = zeros(length(f)))
        P_trace .+= spec.amp .^ 2           # mt_spectrum.amp is the amplitude spectrum
    end

    return f, P_trace
end

"""
    compute_sigma(v, B, rho, dt; NW=4)

Compute normalized cross‐helicity σ_C(f) and normalized residual energy σ_R(f)
as functions of frequency.

Returns
  • f      : frequency vector [Hz]
  • σ_C    : normalized cross‐helicity at each f
  • σ_R    : normalized residual energy at each f
"""
function compute_sigma(v::AbstractMatrix, B::AbstractMatrix, rho::AbstractVector, dt::Real; NW::Real=4)
    # fluctuations
    δv, δb = compute_fluctuations(v, B, rho)

    # trace PSDs
    f, E_v = trace_psd_mt(δv, dt; NW=NW)
    _, E_b = trace_psd_mt(δb, dt; NW=NW)

    # build z± and get their trace PSD
    zplus = δv .+ δb
    zminus = δv .- δb
    _, E_p = trace_psd_mt(zplus, dt; NW=NW)
    _, E_m = trace_psd_mt(zminus, dt; NW=NW)

    # normalized cross‐helicity σ_C = (E₊ – E₋)/(E₊ + E₋)
    σ_C = (E_p .- E_m) ./ (E_p .+ E_m)

    # normalized residual energy   σ_R = (E_v – E_b)/(E_v + E_b)
    σ_R = (E_v .- E_b) ./ (E_v .+ E_b)

    return f, σ_C, σ_R
end
```