### 2025-08-27 notebooks/000_data.qmd
refactor: clean a bit

## Data

```{julia}
using Dates
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using SpacePhysicsMakie: tplot
import SpaceDataModel
using SpaceDataModel: setmeta, times
using TimeseriesUtilities
using Speasy
using GLMakie
using DataFrames, DimensionalData
```

### Parker Solar Probe

- [PSP/FIELDS Level 2 and higher data](https://fields.ssl.berkeley.edu/data/), [SPDF - Coordinated Data Analysis Web (CDAWeb)](https://cdaweb.gsfc.nasa.gov/cgi-bin/eval2.cgi?dataset=PSP_FLD_L2_MAG_RTN&index=sp_phys)

The parameters are:

- 'psp_fld_l2_mag_RTN_4_Sa_per_Cyc' (B in RTN, 0.25s resolution)
- 'psp_swp_spi_sf00_L3_DENS' (Ion density, regular resolution)
- 'psp_swp_spi_af00_L3_DENS' (Ion density, burst resolution)
- 'psp_swp_spi_sf00_L3_VEL_RTN_SUN' (velocity in RTN in Sun's frame, regular resolution)
- 'psp_swp_spi_af00_L3_VEL_RTN_SUN' (velocity in RTN in Sun's frame, burst resolution)
- 'psp_swp_spi_sf00_L3_VEL_RTN' (velocity in RTN in spacecraft frame, regular resolution, you need this to calculate current sheet thickness)
- 'psp_swp_spi_af00_L3_VEL_RTN' (velocity in RTN in spacecraft frame, burst resolution, you need this to calculate current sheet thickness)
- 'Tp_spani_b' (Ion parallel and perpendicular ion temperature, regular resolution. Blue (first element in array): parallel, Red: perp)
- 'Tp_spanib_b' (Ion parallel and perpendicular ion temperature, burst resolution. Blue (first element in array): parallel, Red: perp)

the helium-to-proton abundance ratio ($A_{He} = N_α/N_p × 100\%$, where $N_α$ and $N_p$ are the number densities of helium particles and protons, respectively)

```{julia}
tmin = DateTime("2021-01-15")
n_p = PSP.n_spi(tmin, tmin + Day(1))
n_α = PSP.n_alpha(tmin, tmin + Day(1))

A_He = set_meta(n_α ./ n_p .* 100; ylabel="A_He", units="%", scale=identity)

let tmin = DateTime("2021-01-15")
    tvars = (
        PSP.n_spi,
        PSP.n_alpha,
        A_He
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

Different coordinate frames

```{julia}
tmin = DateTime("2021-01-15")
tvars = [
    spz"cda/PSP_SWP_SPI_SF00_L3_MOM/VEL_RTN_SUN, VEL_SC, VEL_INST, MAGF_SC"...,
    spz"PSP_FLD_L2_MAG_SC_4_SA_PER_CYC/psp_fld_l2_mag_SC_4_Sa_per_Cyc"
]
tplot(tvars, tmin, tmin + Hour(1))
```

"PSP_SWP_SPI_SF00_L3_MOM" ~3.5s resolution

```{julia}
using Discontinuity

tmin = DateTime("2021-01-15")

# Get results from both methods
function get_discontinuity_results(tmin, tmax, B, V)
    @info resolution(B)
    @info resolution(V)
    df_pvi = Discontinuity.ids_finder_PVI(B, V; every=Hour(2))
    df_cvi = Discontinuity.ids_finder_CVI(B, V; every=Hour(2), std_threshold=2)
    df_variance = Discontinuity.ids_finder(B, V; tau=Second(2))
    CVI_ts = CVI(B)
    PVI_ts = PVI(B)
    return (; df_pvi, df_cvi, df_variance, CVI_ts, PVI_ts)
end

function get_discontinuity_results(tmin, tmax, B_prod::Product, V_prod::Product)
    B = DimArray(B_prod(tmin, tmax))
    V = DimArray(V_prod(tmin, tmax))
    return get_discontinuity_results(tmin, tmax, B, V)
end

psp_results = get_discontinuity_results(tmin, tmin + Day(1), PSP.B, PSP.V)
wind_results = get_discontinuity_results(tmin, tmin + Day(1), Wind.B_GSE, Wind.V_GSE_3DP)

# info the number of events
@info "Number of events for PSP" nrow.(getproperty.(Ref(psp_results), (:df_pvi, :df_cvi, :df_variance)))
@info "Number of events for Wind" nrow.(getproperty.(Ref(wind_results), (:df_pvi, :df_cvi, :df_variance)))
```

Test if the CVI and PVI could capture the discontinuities

```{julia}
t0 = DateTime("2021-01-15T09:11")
t1 = DateTime("2021-01-15T09:14")
B_subset = tview(B, t0, t1)
V_subset = tview(V, t0, t1)
df_pvi = Discontinuity.ids_finder_PVI(B_subset, V_subset)
df_cvi = Discontinuity.ids_finder_CVI(B_subset, V_subset)
df_variance = Discontinuity.ids_finder(B_subset; tau=Second(30))

CVI_ts = CVI(B_subset)
PVI_ts = PVI(B_subset)

f = Figure(size=(1400, 1000))

tplot(f[1, 1], [B_subset, [CVI_ts, PVI_ts]])

# Plot 3 candidates from PVI method
for i in 1:min(3, nrow(df_pvi))
    plot_candidate(f[2, 1][i, 1], df_pvi[i, :], [PSP.B,])
end
for i in 1:min(3, nrow(df_cvi))
    plot_candidate(f[2, 2][i, 1], df_cvi[i, :], [PSP.B,])
end
for i in 1:min(3, nrow(df_variance))
    plot_candidate(f[2, 3][i, 1], df_variance[i, :], [PSP.B,])
end
f

```

```{julia}
using Random

f = Figure(size=(1400, 1000))
# Plot 3 candidates from PVI method
for i in 1:min(3, nrow(df_pvi))
    plot_candidate(f[i, 1], df_pvi[i, :], [PSP.B, [CVI_ts, PVI_ts]])
end
random_indices = randperm(nrow(df_cvi))[1:min(3, nrow(df_cvi))]
for (plot_idx, i) in enumerate(random_indices)
    plot_candidate(f[plot_idx, 2], df_cvi[i, :], [PSP.B, [CVI_ts, PVI_ts]])
end
random_indices = randperm(nrow(df_variance))[1:min(3, nrow(df_variance))]
for (plot_idx, i) in enumerate(random_indices)
    plot_candidate(f[plot_idx, 3], df_variance[i, :], [PSP.B, [CVI_ts, PVI_ts]])
end

# Add method labels on the left
Label(f[0, 1], "PVI\nMethod", fontsize=14, tellwidth=false)
Label(f[0, 2], "CVI\nMethod", fontsize=14, tellwidth=false)
Label(f[0, 3], "Variance\nMethod", fontsize=14, tellwidth=false)
resize_to_layout!()
f
```

#### The effect of n / τ on the detection

```{julia}
n_events = 6
n = 16
figsize = (n_events * 300, 400)
f = Figure(; size=figsize)
df = Discontinuity.ids_finder_CVI(B, V; std_threshold=2, n=n)
CVI_tmp = CVI(B; n=n)
# rows = rand(eachrow(df), n_events)

# sort by duration
df = sort(dropmissing(df), :duration; rev=true)
rows = eachrow(df)[1:n_events]
@info rows.duration
for (i, row) in enumerate(rows)
    @info getindex.(Ref(row), (:I_max, :index_std, :index_stability, :index_consistency))
    plot_candidate(f[1, i], row, [PSP.B, CVI_tmp])
end
resize_to_layout!()
f
```

```{julia}
n_events = 6
n = 16
figsize = (n_events * 300, 400)
f = Figure(; size=figsize)
df = Discontinuity.ids_finder(B, V; std_threshold=2, tau=Second(n))

# sort by duration
df = sort(dropmissing(df), :duration; rev=true)
rows = eachrow(df)[1:n_events]
@info rows.duration
for (i, row) in enumerate(rows)
    @info getindex.(Ref(row), (:index_std))
    plot_candidate(f[1, i], row, PSP.B)
end
resize_to_layout!()
f
```

```{julia}
f = Figure(size=(1400, 1000))

for (i, n) in enumerate([1, 2, 4, 32])
    df_cvi = Discontinuity.ids_finder_CVI(B, V; std_threshold=2, n=n)
    CVI_tmp = CVI(B; n=n)
    row = rand(eachrow(df_cvi))
    @info getindex.(Ref(row), (:I_max, :index_std, :index_stability, :index_consistency))
    plot_candidate(f[1, i], row, [PSP.B, CVI_tmp])
    f
end
f
```

```{julia}
f = Figure(size=(1400, 1000))

for (i, n) in enumerate([2, 4, 8, 32])
    df = Discontinuity.ids_finder(B, V; std_threshold=2, tau=Second(n))
    row = rand(eachrow(df))
    @info getindex.(Ref(row), (:index_std, :index_stability))
    plot_candidate(f[1, i], row, [PSP.B])
    f
end
f
```

### THEMIS

Notes: `thb_peim_t3_magQ` is not accurate in solar wind mode.

```{julia}
let tmin = DateTime("2021-01-15")
    tvars = (
        THEMIS.B_GSE,
        THEMIS.B_FGL_GSE,
        THEMIS.n,
    )
    tplot(tvars, tmin, tmin + Day(1))
end
```

### Wind

```{julia}
let tmin = DateTime("2021-01-15"), add_unit = false
    tvars = (
        Wind.V_GSE_K0,
        Wind.V_GSE_3DP,
        [
            SpeasyProduct("WI_K0_SWE/Np"; labels=["Proton"]),
            SpeasyProduct("WI_PM_3DP/P_DENS"; labels=["Proton (3DP, PESA Low)"]),
            SpeasyProduct("WI_H1_SWE/Proton_Np_nonlin", labels=["Proton (non-linear fitting)"]),
            SpeasyProduct("WI_H1_SWE/Alpha_Na_nonlin", labels=["Alpha (non-linear fitting)"]),
            Wind.n_e
        ],
    )
    tplot(tvars, tmin, tmin + Hour(3))
end
```

`Wind.V_GSE_3DP` ("WI_PM_3DP/P_VELS") has higer time resolution than `Wind.V_GSE_K0` ("WI_K0_SWE/V_GSE").

`Wind.n_p_nonlin` and `Wind.n_p_K0` seem to be the same.

![](../figures/others/wind.png)

WI_H1_SWE	Proton_W_moment	        Proton thermal speed W (km/s) from isotropic moment analysis
WI_H1_SWE	Proton_Wperp_moment 	Proton thermal speed Wperpendicular (km/s) from bimax moment analysis
WI_H1_SWE	Proton_Wpar_moment	    Proton thermal speed Wparallel (km/s) from bimax moment analysis

WI_H1_SWE	Proton_Wperp_nonlin	    Proton thermal speed perpendicular to the magnetic field direction [km/s].
WI_H1_SWE	Proton_Wpar_nonlin	    Proton thermal speed parallel to the magnetic field direction [km/s].
WI_PLSP_3DP	Ion omnidirectional fluxes 0.4-3 keV and moments, often at ~24 second resolution, PESA Low, Wind 3DP
WI_PM_3DP	Ion moments (computed on-board) @ 3 second (spin) resolution, PESA LOW, Wind 3DP

### SOLO

"cda/SOLO_L2_MAG-RTN-NORMAL-1-MINUTE/B_RTN",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/N",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/V_RTN",
"cda/SOLO_L2_SWA-PAS-GRND-MOM/T",

### STEREO

"cda/STA_L2_PLA_1DMAX_1MIN/proton_number_density",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_bulk_speed",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_n_s_flow_angle_inst",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_e_w_flow_angle_inst",
"cda/STA_L2_PLA_1DMAX_1MIN/proton_Vr_RTN",