### 2025-10-06 index.qmd
feat: reorganize paper structure with new abstract and discussion sections

---
title: Evolution of solar wind discontinuities in the inner heliosphere
subtitle: PSP and Earth conjunctions and alignments
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    email: vassilis@ucla.edu
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
  - name: Tai Phan
    email: phan@ssl.berkeley.edu
    orcid: 0000-0002-6924-9408
    affiliation: Space Sciences Laboratory, University of California, Berkeley
---

# Abstract

In this study, we analyze three time intervals during which the Parker Solar Probe (PSP) was in close proximity to the Sun. We investigate current sheets (magnetic discontinuities) observed by PSP, as well as by the ARTEMIS and Wind missions.

Using an automated detection method, we identify current sheets in the magnetic field data and examine their properties in conjunction with plasma parameters such as velocity, density, and temperature.

We identified more than 45,000 current sheets across the three selected time intervals, including approximately 17,000 events from the Parker Solar Probe (PSP). Our analysis shows that the absolute current density decreases with increasing radial distance from the Sun, while the normalized current density—defined as the current density divided by the Alfvén current density—remains relatively constant across different heliocentric distances. In contrast, the current sheet thickness increases with radial distance, indicating that current sheets broaden as the solar wind expands. When normalized by the ion inertial length, however, the thickness shows a slight decrease. We also observe distinct distributions in the amplitudes of magnetic field jumps among the different missions. Furthermore, the current sheet thickness and current density are found to be strongly correlated, suggesting that thinner current sheets are associated with stronger current densities.

Most current sheets exhibit only small variations in magnetic field magnitude, suggesting a predominantly force-free, rotational-discontinuity (RD)–like nature. However, significant deviations from the Walén relation are observed, even when pressure anisotropy is taken into account, indicating that anisotropy alone cannot explain the discrepancies expected for ideal RDs. Finally, we find that the Alfvénicity of the current sheets is strongly correlated with the cross helicity and residual energy of the ambient plasma, suggesting that the local current sheet dynamics are largely governed by the surrounding turbulent environment.

<!-- 
In addition, we analyze the Alfvénicity of the current sheets and find that both Parker Solar Probe and Wind observations consistently show that the velocity-jump ratio is often smaller than unity. While the two missions reveal similar statistical trends, the alignment of the magnetic field and velocity jump directions differs, with Parker Solar Probe exhibiting a larger degree of angular misalignment.

In contrast, other parameters such as plasma beta and the alpha-particle abundance do not show a significant influence, and no strong correlations are observed.
-->

# Introduction

Current sheets are ubiquitous mesoscale structures in the solar wind, where the magnetic field changes direction abruptly over ion-scale thicknesses, while their lateral extents can exceed typical magnetohydrodynamic (MHD) scales. These current sheets most readily identified from sharp magnetic-field rotations but can also be evident in simultaneous changes of plasma density, bulk velocity, and temperature.

In turbulent solar wind plasmas, current sheets are widely recognized as signatures of intermittency, contributing to the non-Gaussian nature of magnetic field fluctuations and localized energy dissipation [@borovskyContributionStrongDiscontinuities2010; @grecoPartialVarianceIncrements2017]. They play a crucial role in mediating the transfer of energy across scales, from large-scale MHD fluctuations down to ion and electron kinetic scales, and are believed to be important sites of plasma heating and particle acceleration [@osmanIntermittencyLocalHeating2012; @tesseinAssociationSuprathermalParticles2013].

<!-- Because they concentrate gradients and localized dissipation, current sheets are central features of solar-wind intermittency and non-Gaussian magnetic fluctuations [@borovskyContributionStrongDiscontinuities2010; @grecoPartialVarianceIncrements2017]. They are therefore important for understanding how energy cascades from large MHD scales to ion and electron kinetic scales, and for identifying sites of plasma heating and particle energization [@osmanIntermittencyLocalHeating2012; @tesseinAssociationSuprathermalParticles2013]. -->

Understanding the formation, structure, and evolution of current sheets at different heliocentric distances is therefore essential for elucidating the mechanisms that govern solar wind turbulence and heating. A number of statistical studies have investigated current sheet properties across a broad range of radial distances [@marianiStatisticalStudyMagnetohydrodynamic1983; @artemyevDynamicsIntenseCurrents2018; @perroneCoherentEventsIon2020; @liuCharacteristicsInterplanetaryDiscontinuities2021; @lotekarKineticscaleCurrentSheets2022; @vaskoKineticscaleCurrentSheets2022; @vaskoKineticScaleCurrentSheets2024]. Despite this progress, several important gaps remain.
First, many earlier studies rely on datasets obtained at different times and with instruments of varying temporal resolution, which complicates consistent comparisons across heliocentric distances. Few investigations have examined simultaneous observations from multiple radial positions [@velliUnderstandingOriginsHeliosphere2020, @telloniSpacecraftRadialAlignments2023], limiting our understanding of how current sheet properties evolve as the solar wind expands. Second, existing identification methods often depend on fixed thresholds—such as absolute magnetic-field jumps—or are highly sensitive to data resolution, as in the case of partial variance increment techniques. To enable meaningful cross-mission comparisons, more robust approaches based on normalized, resolution-independent criteria are needed. Finally, most prior work has classified discontinuities within the ideal MHD framework—for instance, distinguishing tangential from rotational discontinuities. While useful in some contexts and extensively studied, these categories become inadequate at ion scales, where kinetic and multi-fluid effects play a central role. In this study, we therefore move beyond strict MHD classifications and instead emphasize the Alfvénic character of current sheets—that is, the extent to which plasma and magnetic-field variations follow Alfvénic correlations [@shenComparingPlasmaAnisotropy2024]. Although this perspective is fundamental for understanding the origin and evolution of current sheets, it has received relatively limited systematic attention.

The Parker Solar Probe (PSP) mission offers an unprecedented opportunity to probe current sheets in the near-Sun environment with high cadence and to compare those observations with near-Earth measurements. In this study, we focus on three PSP perihelion intervals (Encounters 7–9) selected for their differing geometric alignments with the near-Earth spacecraft ARTEMIS and Wind. Encounter 8 features strong radial alignment, enabling sampling of similar solar wind streams while Encounter 7 exhibits weaker alignment and Encounter 9 places PSP roughly 180° away in longitude. These contrasting configurations enable a controlled examination of current sheet properties and, more broadly, provide insight into how those properties depend on the solar wind's source region.

Using a robust automated detection algorithm, we identify and analyze more than 45,000 current sheets across the three intervals (≈17,000 from PSP). We systematically characterize their thickness, current density, magnetic-field jump, and Alfvénicity. The remainder of the paper is organized as follows: Section 2 describes datasets and methods; Section 3 presents results; and Section 4 discusses physical implications and conclusions.

<!-- References

@hudsonDiscontinuitiesAnisotropicPlasma1970

“Knetter [2005] extended this work to intervals in 2002 and 2003 which also included fast wind intervals. He obtained similar results to above and amassed a final total collection of 366 discontinuities. All of the discontinuities had normal field components consistent with zero. He concluded that the discontinuities are best explained as TDs regardless of the amount of dB/B.” ([Vasquez et al., 2007, p. 2](zotero://select/library/items/VPNY4UFZ)) ([pdf](zotero://open-pdf/library/items/DNN2ADMT?page=2&annotation=9Q8CVTE9)) 
-->

# Dataset and method

In this study, we utilize magnetic field and plasma measurements from multiple spacecraft. For the Parker Solar Probe (PSP), we use magnetic field data from the FIELDS instrument and plasma data from the Solar Probe ANalyzer (SPAN) suite. The proton and electron anisotropies are derived from the corresponding SPAN-Ion and SPAN-Electron measurements. For the Wind and ARTEMIS missions, magnetic field data are obtained from the onboard magnetometers, while plasma parameters are taken from the Faraday cup instrument on Wind and the corresponding plasma instrument on ARTEMIS.

Our study focuses on the evolution of solar wind discontinuities within plasma parcels as they undergo expansion. We analyze datasets from the Parker Solar Probe (PSP) and Earth-orbiting missions (e.g., ARTEMIS, DSCOVR, and Wind) at various radial distances in the inner heliosphere, during periods of conjunctions and alignments between PSP and these Earth-orbiting missions.

To achieve this, we utilize a configuration where two spacecraft are radially aligned in the Heliographic (HG) frame, with the longitude fixed on the Sun. It is important to note that the plasma parcels observed by the two spacecraft may differ in composition, as solar wind particles propagate at varying velocities. For instance, protons often exhibit a beam in their velocity distribution function, while the thermal speed of electrons is much higher than their bulk speed. Nonetheless, observing plasma that passes through both spacecraft offers the best opportunity to study the evolution of solar wind discontinuities with minimal influence from other processes during propagation. It is also expected that these plasma parcels evolve due to heating and other processes.

<!-- ::: {#fig-plasma-adiabatic-evolution .column-page}
![](figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR-next.svg)

![](figures/evolution/plasma-adiabatic-evolution_e2-DSCOVR.svg)

Comparison of plasma dimensionless quantities (plasma beta and normalized plasma speed by Alfven speed) between PSP (red) and DSCOVR (blue). Yellow dots are predicted plasma beta and normalized speed using formulary from @perroneRadialEvolutionSolar2019.
::: -->

The plasma parcels in question are large-scale structures. As the scale size decreases below the mesoscale [@viallMesoscaleStructureSolar2021] (with radial sizes ranging from approximately 5 × 10³ km to 10⁷ km), it becomes increasingly unlikely to identify the same structure, depending on the latitude difference between the two spacecraft. Furthermore, although large-scale structures (especially density enhancements) can be identified as passing through both spacecraft [@berriotIdentificationSinglePlasma2024], it is nearly impossible to detect kinetic-scale structures, such as solar wind discontinuities (with time scales on the order of seconds), by both spacecraft simultaneously. Therefore, our study focuses on the statistical properties of solar wind discontinuities within the "same" large-scale structures that pass through both spacecraft.

After identifying the periods of conjunctions and alignments between PSP and Earth-orbiting missions, we apply the methodology described in [@???] to identify and characterize solar wind discontinuities. Generally, discontinuities observed by PSP are shorter (1-5 seconds) and exhibit higher current density. Consequently, we use different time windows for identifying discontinuities: 15 seconds for PSP and 30 seconds for Earth-orbiting missions. A example of the discontinuities observed by PSP is shown in @fig-examples.

::: {#fig-examples .column-page}

![](figures/examples/psp_e2_example.png)

Solar wind discontinuities examples observed by PSP fitted by a logistic function in LMN coordinates.
:::

<!-- 
PSP:
SWEAP consists of four separate sensors that measure the ions and electrons that make up the plasma in the solar wind and solar corona. The Solar Probe ANalyzer (SPAN) instruments are electrostatic analyzers (ESAs) that use curved plates to measure the 3-D particle distribution from within the shadow cast by the spacecraft thermal shield. Two of the sensors are co-located on the ram side of the spacecraft: SPAN-Ai measures ions, and SPAN-Ae measures electrons. SPAN-B measures electrons and is on the anti-ram side of the spacecraft. The Solar Probe Cup (SPC) is a Faraday Cup instrument that is directly illuminated by the Sun and measures both ions and electrons.

"Looking at your intervals, which occurred during the early phase of the mission when PSP was far from the Sun, I realize that the plasma data wasn't good for your purpose because (1) SPAN-ion during at least the first 3-4 orbits did not measure the solar wind well, since a good portion of the solar wind distributions was out of view of that instrument, and (2) data from SPC (Farady Cup) does not give good temperature data. I am wondering if you have conjunction events that occurred later, say starting Encounter 7, when data from SPAN would be accurate. If not, some parts of Encounters 4-6 might be OK."
-->

For this study

We utilized high-resolution magnetic field and plasma data from the FIELDS and SWEAP instruments aboard the Parker Solar Probe (PSP). For the plasma data presented in the figures of this paper, we used the dataset from SPAN-i, which measures ions in the solar wind. Although ion data is also available from SPC, the SPAN-i data was chosen due to its higher temporal resolution, making it more suitable for analyzing discontinuities. Our analysis confirms that the results obtained from SPC are consistent with those derived from SPAN-i data. It is also worth noting that sqtn data from the Radio Frequency Spectrometer (RFS) and Low-Frequency Receiver (LFR) for density measurements is available. However, we generally find that the results obtained from RFS/LFR do not differ significantly from those derived from SPAN-i data. Due to the limitations of SPAN-i data during the early mission phases, when a significant portion of the solar wind distributions was out of the instrument’s view, our study concentrates on data beginning with Encounter 7, where SPAN-i provides more accurate measurements.

<!-- 
ARTEMIS:
We use the following data collected by ARTEMIS P1: magnetic fields with a ∼4s spin resolution (Auster et al. 2008), the ion bulk velocity v, and the ion (assumed to be proton) temperature (scalar) Tp calculated from velocity distributions measured at spin resolution by the Electrostatic Analyzer (ESA; McFadden et al. 2008). Being designed for measurements of a hot, rarefied plasma in Earth’s magnetosphere, ESA typically overestimates the proton temperature in the solar wind, because the solar wind proton beam is narrower than ESA’s energy and angular resolution.
TODO: check if the ESA is in solar wind mode
-->

We utilized data collected by ARTEMIS, including magnetic fields measured at approximately 5 Hz resolution by the Fluxgate Magnetometer [@austerTHEMISFluxgateMagnetometer2008], as well as plasma velocity ($v$) and density ($n$), which were derived from velocity distribution data obtained by the Electrostatic Analyzer with a resolution of around 4 seconds [@mcfaddenTHEMISESAPlasma2009]. It is important to note that the ESA tends to overestimate proton temperature in the solar wind due to the narrower proton beam compared to the ESA’s energy and angular resolution. Therefore, any findings related to temperature anisotropy should be interpreted with caution.

<!-- 
WIND:
Wind 3dp, PESA LOW 1 spin resolution ion (proton and alpha) moments
Wind 3D Plasma Analyzer
Instrument Description
The Wind 3DP instrument consists of six different sensors. There are two electron (EESA) and two ion (PESA) electrostatic analyzers with different geometrical factors and field-of-views covering the energy range from 3 eV to 30 keV. There are also a pair of solid state telescopes (SST) that measure electrons with energies up to 400 keV and protons with energie
-->

The following data collected by Wind are used: magnetic fields with a $\sim 11$ Hz resolution measured by the Magnetic Field Investigation (MFI) [@leppingWINDMagneticField1995], the ion (proton and alpha) moments measured by ion (PESA) electrostatic analyzers of the 3D Plasma Analyzer (3DP) [@linThreedimensionalPlasmaEnergetic1995].

> We will also present plasma density estimates delivered by the quasi-thermal noise spectroscopy available at about 7 s cadence (Moncuquet et al. 2020).
> The procedure of CS selection and methodology, briefly described below, is equivalent to those used in the paper

@telloniSpacecraftRadialAlignments2023

<!-- sensing solar wind emitted from the same region on the solar surface, with the caveat that dynamical processes at the chromosphere–corona interface have not dramatically altered the solar footpoint within the time corresponding to the radial propagation from the measurement of one spacecraft to that of the other. -->

## Methodology

To identify current sheets, we employ an automated detection algorithm similar to that used in our previous study (Zhang et al., 2024). The magnetic field data are divided into consecutive time intervals, and the standard deviation within each interval is calculated and compared with those of adjacent intervals. A current sheet is identified when the standard deviation within an interval significantly exceeds that of its neighboring intervals, indicating a sharp magnetic field variation.

Once the current sheets are identified, we apply minimum variance analysis (MVA) to transform the magnetic field and plasma data into the local LMN coordinate system, where L, M, and N correspond to the maximum, intermediate, and minimum variance directions, respectively. The spatial thickness of each current sheet is estimated by multiplying its temporal duration by the plasma velocity projected along the normal (N) direction. The current density is then derived accordingly.

For each identified event, we select time steps immediately before and after the current sheet as the upstream and downstream intervals, respectively. The mean plasma density and temperature from these intervals are used to estimate the plasma pressure and pressure anisotropy factors. Using these parameters, we derive both the Alfvén velocity and a modified Alfvén velocity that accounts for the pressure anisotropy. Finally, we compute the cross helicity and residual energy within the same intervals to characterize the local solar wind conditions surrounding each current sheet.

# Overview and Example of Current Sheets

> @lotekarKineticscaleCurrentSheets2022 Figure 1 presents an overview of the considered 10 day interval. Over this interval, the spacecraft was at radial distances from 0.17 to 0.24 au, that is, from 35 to 50 solar radii, and remained within the inward magnetic field sector without crossing the heliospheric current sheet (Phan et al. 2020; Szabo et al. 2020). No coronal mass ejections were observed in the considered interval (Phan et al. 2020; Szabo et al. 2020). The spacecraft was approximately corotating with the Sun and connected to the same coronal hole (Bale et al. 2019; Badman et al. 2020). Panels (a)–(d) present 1 minute averages of the magnetic field magnitude, proton flow velocity, densities, and temperatures of protons and electrons as well as plasma density estimates provided by the quasi-thermal noise spectroscopy. The magnetic field magnitude was about 100 nT at the perihelion on November 6, and around 50 nT as the spacecraft moved to a radial distance of 50 solar radii from the Sun on November 1 and 10. The solar wind was typically slow  with a proton flow velocity below 400 km s−1, except for the ≈20 hr period of fast solar wind between November 9 and 10. The three plasma density estimates in panel (c) are consistent with each other within a few tens of percent over the entire interval, except for the fast solar wind interval, where the proton densities were about 4 times smaller than the plasma density estimates provided by the quasi-thermal noise spectroscopy. We will use the electron density estimates, calibrated to best match proton densities and the results of quasi-thermal noise spectroscopy, as plasma density estimates. The proton temperature during the considered interval varied between about 10 and 100 eV, while the electron temperature remained around 30 eV. Panel (e) presents 1 minute averages of proton and electron betas and shows that both quantities varied in the range from 0.1 to 5.
> Figure 2 presents several CSs from our data set.

## Examples of discontinuities

![](figures/event-example.pdf){#fig-event-example}

<!-- ![](figures/examples/psp_e2_example.png) -->

In general discontinuities observed by PSP are with

- shorter duration (1~5 secs)

  compared with 5-10 secs

- large current density (100-1000 nA/m^2)

  compared with 1-10 nA/m^2

# Results

We analyze the Alfvénicity of the current sheets and find that both Parker Solar Probe and Wind observations consistently show that the velocity-jump ratio is often smaller than unity. While the two missions reveal similar statistical trends, the alignment of the magnetic field and velocity jump directions differs, with Parker Solar Probe exhibiting a larger degree of angular misalignment.

Figure 1 presents the statistical distributions of current density, current sheet scale (or cycle length), and magnetic field jumps derived from different missions and time intervals. Despite the varying observation conditions, the normalized current density exhibits similar magnitudes across the Parker Solar Probe (PSP), Wind, and ARTEMIS datasets. In contrast, the current sheet scale increases systematically from PSP to ARTEMIS, suggesting a radial evolution as the solar wind expands. However, after normalization by local parameters, the current sheet scale shows a slight decrease.

The magnetic field jump displays more complex behavior among the missions, with PSP, Wind, and ARTEMIS showing distinct distributions. This diversity may indicate the presence of two populations of current sheets in the near-Sun environment, possibly reflecting different formation mechanisms or solar wind regimes.

Figure 2 shows the joint distributions between (a) current density and current sheet scale, (b) current density and magnetic field jump, and (c) current sheet scale and magnetic field jump. A clear negative correlation is observed between current density and current sheet scale, indicating that stronger current densities are generally associated with thinner current sheets.

![](figures/overview-7-all.pdf){#fig-overview}

![](figures/properties_hist-mva.pdf){#fig-properties-hist}

Figure 3 illustrates the parameters used to characterize the Alfvénicity of the current sheets. Two key quantities are considered: (1) the velocity jump across the current sheet, and (2) the variation in magnetic field magnitude across the same interval. The first panel presents the ratio of magnetic field magnitude before and after the current sheet, while the second panel shows the angular difference between the Alfvén velocity and the plasma velocity.

In general, the magnetic field magnitude ratio is less than unity, with mean values of approximately 0.6 for the Parker Solar Probe and 0.5 for the Wind observations. The velocity angle distribution is concentrated near 0° and 180°, indicating that the plasma and Alfvén velocities are typically aligned or anti-aligned. The Wind data exhibit a somewhat broader angular distribution compared to PSP, suggesting a larger degree of deviation from ideal Alfvénic alignment. The variation in magnetic field magnitude across the current sheets is relatively small, implying that the magnetic field strength remains nearly constant while its direction rotates—consistent with the rotational nature of many observed current sheets.

![](figures/joint_properties-mva.pdf){#fig-joint-properties}

![](figures/Alfvenicities.pdf)

To further investigate how the Alfvénicity of current sheets relates to other parameters, we combine the velocity jump ratio and the velocity jump angle into a single scalar parameter, Q, following the formulation introduced by Sonnerup (2016). The parameter Q, bounded between –1 and 1, quantifies the deviation from ideal Alfvénic behavior expected for a rotational discontinuity.

We examine the relationship between Q and various plasma properties and find that most parameters show little or no correlation. However, two parameters exhibit strong correlations with Q: the cross helicity and the residual energy of the surrounding solar wind. Both quantities describe the Alfvénicity of the background solar wind rather than that of individual current sheets. Our results show that current sheet Alfvénicity increases with solar wind cross helicity and decreases with residual energy.

Additionally, for the Parker Solar Probe observations, we find a weak positive correlation between the temporal duration of the current sheets and their Alfvénicity, with longer-duration current sheets tending to display slightly higher Alfvénicity. However, this trend may partly arise from instrumental limitations, as longer-duration events provide better temporal resolution for plasma measurements.

![](figures/Q_sonnerup_joint_dist_den.pdf)

We also analyze the rotational characteristics of the current sheets and their relationship with the normal component of the magnetic field. The rotational profile provides insights into how the magnetic field vector changes direction across each current sheet. Our results show that the Parker Solar Probe observes a distinctly different distribution of rotation angles compared to the Wind and ARTEMIS missions. This discrepancy suggests that the orientation and rotational structure of current sheets evolve significantly as the solar wind propagates from the near-Sun environment to near-Earth distances.

![](figures/B_n_ω_subset=true.pdf)

Comparison between Alfven speed and plasma speed change

For rotational discontinuities, the plasma velocity jump across them: the plasma flow velocity $v_l$ (the solar wind velocity projected onto $l$) changes due to changes in $B_l$, $Δv_l=±Δv_A$ with $v_A= B_l / \sqrt{4 n m}$.

![](figures/enc7/dvl.png)

K2: $Δv_l$,l demonstrates a strong correlation with $Δv_{A}$, albeit being consistently smaller.

## The effect of alignment on discontinuity properties

![](figures/thickness_current.png)

![](figures/B_angle.png)

# Discussion

K1: Normalized discontinuities properties does not vary much with radial distance

Thickness and current density relationship, scale-dependent properties of current sheets in the solar wind

Discussion

Our analysis reveals several robust, physically meaningful trends that illuminate how current sheets evolve from the near-Sun environment sampled by Parker Solar Probe (PSP) to the near-Earth environment sampled by Wind and ARTEMIS. Below I synthesize the principal findings, interpret their physical significance, identify methodological caveats, and suggest targeted follow-ups.

1. Alfvénicity is controlled by the background turbulence, not only by local sheet properties.
The strong positive correlation between current-sheet Alfvénicity and the ambient cross-helicity, together with the negative correlation with residual energy, indicates that the degree to which a current sheet behaves like an ideal rotational discontinuity is largely set by the surrounding turbulence state. High cross-helicity (more Alfvénic background fluctuations) promotes velocity–magnetic-field alignment at the sheet and hence higher sheet Alfvénicity; conversely, large (negative) residual energy — a relative excess of magnetic over kinetic fluctuation energy — corresponds to less Alfvénic sheets. This pattern is consistent with a picture in which current sheets are embedded, and shaped, by the local turbulent cascade rather than being isolated structures created by a single local process. In other words, the macroscopic turbulence environment biases whether a boundary will appear Alfvénic or non-Alfvénic when sampled by a spacecraft.

2. Stronger currents ↔ thinner sheets: intermittency and cascade interpretation.
The pronounced anti-correlation between current density and sheet scale (cycle length) — stronger current density associated with smaller spatial thickness — is a hallmark of intermittency in MHD turbulence. In a turbulent cascade, gradients steepen locally and concentrate current into thin, intense sheets. That we observe similar normalized current densities across radial distances suggests the cascade produces structures that are, after normalization by local parameters (e.g., Alfvén speed), roughly self-similar. The increase of absolute cycle length with heliocentric distance (and the slight decrease after normalization) can be interpreted as geometric expansion and evolution of structures during propagation combined with changes in local plasma conditions.

3. Two populations and the predominance of rotational behavior.
The magnetic magnitude change across many sheets is small while the direction rotates, consistent with a predominance of rotational discontinuities (RDs). Yet the mission-dependent diversity in magnetic jump distributions and the suggestion of two populations (one with small |ΔB| and one with larger |ΔB|) point to coexisting RD-like and more tangential/discontinuous (TD-like) structures near the Sun. PSP’s distinct rotation-angle distribution relative to Wind/ARTEMIS indicates evolution: as the solar wind propagates, some RD-like structures may transform (via interactions, steepening, or merging) into structures with different jump statistics, or different source populations (e.g., switchback boundaries versus streamer-belt tangential discontinuities) could dominate at different radii.

4. Why pressure anisotropy alone fails to explain the velocity vs. magnetic jumps.
Our finding that pressure anisotropy does not reconcile the observed discrepancy between the magnetic-field-derived Alfvén velocity jump and the measured plasma velocity jump suggests that single-fluid MHD (even with scalar anisotropy corrections) is insufficient. Possible contributors include: multi-fluid effects (e.g., decoupling of alpha particles or beams from the proton bulk), Hall and finite-Larmor-radius physics at ion scales, kinetic pressure tensor terms, or wave/beam populations that modify the bulk velocity without a matching magnetic signature. This motivates direct inspection of ion distribution functions (alpha particles, proton beams), search for Hall or reconnection signatures, and targeted kinetic modeling (hybrid/PIC) to quantify non-MHD contributions.

5. Temporal-duration effect: physical or instrumental?
The weak positive relation between current-sheet duration and Alfvénicity (longer events appearing more Alfvénic) must be interpreted cautiously. Longer durations provide more plasma samples and therefore better statistical estimates of upstream/downstream velocities; thus part of the trend may be instrumental or sampling bias. Nevertheless, if a residual physical trend remains after accounting for resolution effects, it could indicate that more spatially extended sheets preserve the Alfvénic coupling between flows and fields more effectively than shorter, more transient sheets — again pointing to an intimate connection between sheet morphology and the ambient turbulence.

6. Radial evolution, rotation profiles, and implications for solar-wind dynamics.
The systematic differences in rotation-angle distributions (PSP versus Wind/ARTEMIS) suggest that current sheets do not simply convect unchanged: they undergo reconfiguration. Processes that can drive this evolution include differential expansion, interactions with other structures (compressive regions, shocks, stream interactions), and nonlinear turbulent evolution (coalescence, steepening). This evolution has practical implications: the dynamical role of current sheets in particle energization, wave generation, and transport will depend on location and on how their topology and Alfvénicity change with distance.

7. Methodological caveats and uncertainties.
Several limitations could affect quantitative interpretation and should guide follow-up analyses: (i) single-spacecraft MVA and thickness estimates carry uncertainties in the normal direction and the projection of velocity; (ii) current density estimated from ΔB/Δl assumes a 1D geometry and neglects possible multi-dimensional structure; (iii) time-resolution differences between magnetometers and plasma instruments (SPAN, Faraday cup) can bias velocity-jump estimates; and (iv) detection thresholds and interval selection in the automated algorithm can preferentially select certain sheet types. We therefore recommend sensitivity tests varying detection thresholds, bootstrapping MVA normals, and quantifying the effect of cadence by sub-sampling or synthetic down-sampling tests.

8. Specific, actionable follow-ups.
To strengthen the physical interpretation and disentangle competing hypotheses, we recommend:
	•	Cross-checking events with high-cadence particle moments and full distribution functions (to examine alpha/beams and non-Maxwellian features).
	•	Searching for localized signatures of Hall physics or reconnection (bipolar electron jets, outflow pairs, Hall magnetic-field patterns) in high-resolution fields and particle data.
	•	Conditioning the statistics on solar-wind type (fast vs. slow, presence of switchbacks, stream origin) to see whether the two inferred populations map to flow regimes.
	•	Running hybrid or kinetic simulations initialized with observed upstream parameters to test whether pressure-tensor or multi-species effects reproduce the velocity–field jump mismatch.
	•	Expanding the radial coverage (include Solar Orbiter, STEREO where possible) or employing multi-point analysis to reduce geometric ambiguities.

9. Concluding synthesis.
Taken together, our results support a picture in which current sheets are predominantly turbulence-embedded, dynamically evolving structures whose Alfvénic character is governed more by the ambient solar-wind turbulence state (cross-helicity and residual energy) than by simple local pressure anisotropy. The tight coupling between sheet intensity and spatial scale, the presence of (at least) two structural populations, and the radial evolution of rotation profiles together point to a complex interplay of cascade intermittency, expansion effects, and kinetic physics. Resolving the remaining discrepancies will require higher-resolution particle data, multi-spacecraft constraints, and targeted kinetic modeling.

# Conclusion

- The properties of the discontinuities is related to the local plasma parameters
  - Thickness => ion inertial length
  - Current density => Alfven velocity (current density)
- Normalized thickness and current density of discontinuities remain constant with radial distance
- Anisotropy of the plasma is expected to be larger near the sun to explain the observed speed change ratio $Δv_i/Δv_A$.

- Normalized thickness and current density of discontinuities remain constant with radial distance (negligible change compared to their spread). In contrast to @artemyevDynamicsIntenseCurrents2018 results that show a decrease in normalized thickness and increase in normalized current density with radial distance.

- Better alignment period has a slightly better agreement of the properties of the discontinuities (normalized thickness, current density, $|\Delta \mathbf{B}/B|$ and rotation angle) between PSP and Earth-orbiting missions.
- $B_N/B$ and in-plane rotation angle are significantly different between PSP and Earth-orbiting missions independent of the alignment period.

# References

-   @velliUnderstandingOriginsHeliosphere2020

-   @artemyevDynamicsIntenseCurrents2018 examined the evolution of transient currents in the solar wind using a data set compiled from observations of the same solar wind ﬂow at Earth’s and Mars's orbits.

    ```         
    > we show that it consists of several processes: discontinuity thinning (decrease in thickness normalized by the ion inertial length), intensification of currents normalized to the proton thermal current (i.e., the product of proton charge, density, and thermal velocity), and increase in the compressional component of magnetic ﬁeld variations across discontinuities.
    ```

# Appendix