### 2025-10-29 index.qmd
docs: methods (#1)

See [article.qmd](article.qmd) for the main article.

See [notebooks/00_index.qmd](notebooks/00_index.qmd), [notebooks/02_stats.qmd](notebooks/02_stats.qmd), and [notebooks/10_results.qmd](notebooks/10_results.qmd) for the data analysis.

See [github](https://github.com/beforerr/psp_conjunction) for the code.