### 2025-10-29 notebooks/00_index.qmd

Interactively exploring the quantities of solar wind. See [Data](000_data.qmd) for more details about different data products.

## Setup

```{julia}
using Dates
using Speasy
using DrWatson
using PSPAnisotropy
using PSPAnisotropy: Wind, PSP, THEMIS
using Discontinuity
using Discontinuity: compute_params!, filter_low_mva_quality
using Discontinuity: compute_Alfvenicity_params!, Alfven_velocity
using SPEDAS
using DimensionalData
using DataFrames, DataFramesMeta
using SpacePhysicsMakie, GLMakie
using SPEDAS: setmeta, setmeta!
using Beforerr: _theme_legend, easy_save, hideylabel!, align_ylabels!
using PartialFunctions

𝐧 = :mva
𝒎 = var_mapping(; 𝐧)

using Unitful
Unitful.preferunits(u"nA", u"km")
unitify(u) = (*)$u
unitify(u::String) = (*)$uparse(u)
includet("../src/plot.jl")

set_Z_theme!()
```

```{julia}
df = workload()

psp_df = @rsubset(df, :id == "PSP"; view=true)
wind_df = @rsubset(df, :id == "Wind"; view=true)
thm_df = @rsubset(df, :id == "THEMIS"; view=true)
```

```{julia}
compute_params!(df)
compute_Alfvenicity_params!(df)

compute_PSP_anisotropy_params!(psp_df)
for enc in 7:9
    tmin_wind, tmax_wind = get_timerange(enc)[2]
    wind_df_enc = @rsubset(wind_df, :enc == enc; view=true)
    compute_Wind_anisotropy_params!(wind_df_enc, tmin_wind, tmax_wind)

    thm_df_enc = @rsubset(thm_df, :enc == enc; view=true)
    compute_THM_anisotropy_params!(thm_df_enc, tmin_wind, tmax_wind)
end
```

## Event example (simple)

```{julia}
let add_fit = true, add_Va = true
    f = Figure(; size=(600, 400))
    psp_row = filter_low_mva_quality(psp_df)[20, :]
    faxs, = plot_candidate(f[1, 1], psp_row, psp_conf, Second(6); add_fit, legend=nothing, add_Va)
    align_ylabels!(faxs.axes)
    thm_row = filter_low_mva_quality(thm_df)[18, :]
    faxs, = plot_candidate(f[1, 2], thm_row, thm_conf, Second(8); add_fit, add_Va)
    hideylabel!.(faxs.axes)
    add_labels!(f[1, 1]; position=Left(), padding=(30, -30, 0, 0), labels=1:3, open="(a.")
    add_labels!(f[1, 2]; position=Left(), padding=(30, -30, 0, 0), labels=1:3, open="(b.")
    # faxs, = plot_candidate(f[1, 3], filter_low_mva_quality(wind_df)[2, :], wind_conf, Second(6); add_fit)
    # hideylabel!.(faxs.axes)

    @info "PSP" psp_row.ΔV_ratio psp_row.ΔV_cosθ
    @info "THEMIS" thm_row.ΔV_ratio thm_row.ΔV_cosθ

    colgap!(f.layout, 1)
    resize_to_layout!(f)
    easy_save("event-example")
    f
end
```

![](../figures/event-example.png)

┌ Info: PSP
│   psp_row.ΔV_ratio = 0.31703819329006777
└   psp_row.ΔV_cosθ = -0.8208860063945036
┌ Info: THEMIS
│   thm_row.ΔV_ratio = 0.3268322029115623
└   thm_row.ΔV_cosθ = -0.9457794099717979

```txt
  :t_us              => 2021-01-15T04:05:04.549
  :t_ds              => 2021-01-15T04:05:17.957
  :tstart            => 2021-01-15T04:05:04
```

## Overview

```{julia}
using Speasy
using SPEDAS: tnorm_combine, DataSet
using DimensionalData
# using CairoMakie
using GLMakie
using Statistics
using PartialFunctions
using Unitful
using SPEDAS.SpaceDataModel: set, extend, @set

import AlgebraOfGraphics as AoG

```

```{julia}
mva_transform((V, B), tmin, tmax; tb_min=tmin, tb_max=tmax) = mva(V(tmin, tmax), B(tb_min, tb_max))
```

```{julia}
tmin_psp = DateTime("2021-01-15")
tmax_psp = DateTime("2021-01-20")
tmin_wind = DateTime("2021-01-17")
tmax_wind = DateTime("2021-01-23")
_tmax_wind = DateTime("2021-01-15T02")

# tmin_psp = DateTime("2021-01-16")
# tmax_psp = DateTime("2021-01-19")
# tmin_wind = DateTime("2021-01-17")
# tmax_wind = DateTime("2021-01-18")
```

### Alfvén ratio over time

```{julia}
f = Figure(; size=(800, 600))
faxs = plot_PSP(f[1, 1], tmin_psp + Day(2), tmin_psp + Day(3), get_vl_ratio_ts(psp_df)...)
modify_faxs_base!(faxs)
# modify_faxs_vl!(faxs)
f
```

```{julia}
let tmin = tmin_wind, tmax = tmin_wind + Day(3)
    f = Figure(; size=(800, 600))
    faxs = plot_Wind(f[1, 1], tmin, tmax, get_vl_ratio_ts(wind_df)...)
    modify_faxs_base!(faxs)
    f
end
```

```{julia}
using Beforerr: add_labels!, align_ylabels!
update_theme!(; Lines=(linewidth=1, alpha=0.9), figure_padding=(1, 1, 1, 15))

function add_wind_twin!(gl, wind_axs)
    # Create a twin axis for Wind magnetic field and density (multiply by the radial ratio)
    # 1AU / 1Rsun =215
    B_ax = wind_axs.axes[1]
    B_ax2 = Axis(gl[1, 1]; yaxisposition=:right, rightspinevisible=true, rightspinecolor=:darkgray)
    hidexdecorations!(B_ax2, grid=false)
    ylims!(B_ax, 0.1, 9.9)
    ylims!(B_ax2, 0.1 * (0.1)^(-1.59), 9.9 * (0.1)^(-1.59))
    n_ax = wind_axs.axes[2]
    n_ax2 = Axis(gl[2, 1]; yaxisposition=:right, rightspinevisible=true, rightspinecolor=:darkgray)
    hidexdecorations!(n_ax2, grid=false)
    ylims!(n_ax, 3, 16)
    ylims!(n_ax2, 3 * (0.1)^(-1.96), 16 * (0.1)^(-1.96))222
end

function overview_plot(enc)
    (tmin_psp, tmax_psp), (tmin_wind, tmax_wind) = get_timerange(enc)
    legend = (; position=Top())

    f = Figure(; size=(1000, 800))
    psp_axs = plot_PSP(f[1, 1], tmin_psp, tmax_psp, get_vl_ratio_ts(psp_df)...; legend)
    modify_faxs_base!(psp_axs)
    gl = GridLayout(f[1, 2])
    wind_axs = plot_Wind(f[1, 2], tmin_wind, tmax_wind - Day(1), get_vl_ratio_ts(wind_df)...; legend)
    modify_faxs_base!(wind_axs)

    labels = (; labels='1':'9', position=Left(), padding=(32, -32, 0, 0))
    add_labels!(f[1, 1]; labels..., open="(a.")
    add_labels!(f[1, 2]; labels..., open="(b.")
    colgap!(f.layout, 2)

    hideylabel!.(wind_axs.axes)
    add_wind_twin!(gl, wind_axs)
    align_ylabels!(psp_axs.axes)
    easy_save("overview-$enc-all")
    return f
end

overview_plot(7)
# overview_plot.(7:9)
```

![](../figures/overview-7-all.png)
![](../figures/overview-8-all.png)
![](../figures/overview-9-all.png)

## Event example (with pressure anisotropy)

```{julia}
using Dates
using Discontinuity: ids_finder
tau = Second(30)
psp_events = ids_finder(PSP.B, tmin_psp, tmax_psp, tau)
wind_events = ids_finder(Wind.B_GSE, tmin_wind, tmax_wind, tau, Wind.V_GSE, Wind.n_p)
thm_events = ids_finder(THEMIS.B_FGL_GSE, tmin_wind, tmax_wind, tau)
```

```{julia}
psp_B_mva = set(mva ∘ PSP.B; name="mva_B", labels=𝑳.B_LMN)
psp_V_mva = Product((PSP.V, PSP.B), mva_transform; ylabel=𝒀.V, labels=𝑳.V_LMN)
psp_eTemp_ani = Product(read_electron_temperature(), tview; labels=["Electron (parallel)", "Electron (perpendicular)"])
psp_pTemp_ani = Product(read_proton_temperature(), tview; labels=["Proton (parallel)", "Proton (perpendicular)"])
psp_Temp_ani = [psp_eTemp_ani, psp_pTemp_ani]
psp_mva_all = [psp_B_mva, psp_V_mva, psp_Temp_ani]

thm_B_mva = set(mva ∘ THEMIS.B_FGL_GSE; name="mva_B", labels=𝑳.B_LMN)
thm_V_mva = Product((THEMIS.V_GSE, THEMIS.B_FGL_GSE), mva_transform; ylabel=𝒀.V, labels=𝑳.V_LMN)
thm_mva_all = [thm_B_mva, thm_V_mva, THEMIS.T_ani]
```

```{julia}
tvars2_wind = [
    tnorm_combine ∘ Wind.B,
    Wind.n,
    Wind.V_GSE,
    Wind.Temp,
    Wind.Temp_ani
]

tvars_themis2 = [[THEMIS.B_FGL_GSE, THEMIS.n, THEMIS.V_GSE, THEMIS.T]; thm_mva_all]
```

```{julia}
f = let
    f = Figure(; size=(1200, 800))
    # plot_candidate(f[1, 1], Event(psp_events, 2), [tvars_psp; psp_mva_all]; link_xaxes)

    wind_event = AniEvent(wind_events, 1) # 300
    wind_T3 = extend(Wind.pTemp_ani, Second(20))
    wind_tvars = make_mva_products(wind_event, Wind.B_GSE, Wind.V_GSE, Wind.n_p, wind_T3)
    wind_tvars = [wind_tvars[3:end]; Wind.Temp_ani; Wind.n; Wind.Temp]
    for p in Wind.Temp_ani
        @info "$p" data = p(wind_event)
    end
    wind_plots = plot_candidate(f, wind_event, wind_tvars, tau)
    # themis_plots = plot_candidate(f[1, 3], Event(thm_events, 3), tvars_themis2, Minute(1); link_xaxes)
    # hideydecorations!(wind_plots)
    # hideydecorations!(themis_plots)
end
```

## Batch processing

Wind Speed, Time to Earth
300 km/s, ~5.8 days
400 km/s, ~3.9 days
500 km/s, ~3.5 days
700 km/s, ~2.5 days

## The effect of tau

```{julia}
psp_df_taus = map([2, 4, 8, 16, 32]) do tau
    psp_conf["tau"] = tau
    df = PSPAnisotropy.produce(psp_conf, psp_conf["timeranges"][1]...)
    @transform! df :tau = tau
    compute_Alfvenicity_params!(df)
end

wind_df_taus = map([2, 4, 8, 16, 32]) do tau
    wind_conf["tau"] = tau
    df = PSPAnisotropy.produce(wind_conf, wind_conf["timeranges"][1]...)
    @transform! df :tau = tau
    compute_Alfvenicity_params!(df)
end
```

Check examples

```{julia}
f = Figure(; size=(1200, 800))
# let df = psp_df_taus[1], config = psp_conf
begin
    df = @subset(wind_df_taus[1], :index_std .> 4)
    config = wind_conf
    event = filter_low_mva_quality(df)[5, :]
    fax, res_psp = plot_candidate(f[1, 1], event, config, Second(10))
    plot_var_info!(fax, event)
    @info "v_l_ratio" event.V_l_ratio event.V_l_ratio_max
end
f
```

```{julia}
f = Figure()
axR1 = Axis(f[1, 1])
axn1 = Axis(f[2, 1])
filter_func = df -> @subset(df, :index_std .> 2)

for (i, df) in enumerate(psp_df_taus)
    v_l_ratio_ts, n = get_vl_ratio_ts(filter_func(df), :V_l_ratio_max)
    lines!(axR1, v_l_ratio_ts)
    lines!(axn1, n)
end

axR2 = Axis(f[1, 2])
axn2 = Axis(f[2, 2])
for (i, df) in enumerate(wind_df_taus)
    v_l_ratio_ts, n = get_vl_ratio_ts(filter_func(df), :V_l_ratio_max)
    lines!(axR2, v_l_ratio_ts)
    lines!(axn2, n)
end

ylims!.((axR1, axR2), 0, 0.8)
# ylims!.((axn1, axn2), 0, 121)
easy_save("vl_ratio-tau")
f
```